---
title: "Size symbolism in production and perception"
subtitle: "A Bayesian analysis of duration, pitch, and animacy"
author:
  - Alexandra Schmidt
  - Aleksandra Ćwiek
  - Petra Wagner
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
bibliography: MAIconicity.bib
link-citations: true
editor_options:
  chunk_output_type: console
---

```{r source setup, echo = TRUE, message=FALSE, warning = FALSE}
########## folders ##########
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
parentfolder  <- dirname(getwd())
models        <- paste0(parentfolder, '/models/')
plots         <- paste0(parentfolder, '/plots/')
scripts       <- paste0(parentfolder, '/scripts/')
data_dir      <- paste0(parentfolder, '/data/')   # adjust if data lives elsewhere

########## folder creation (security: only if missing) ##########
if (!dir.exists(plots))   dir.create(plots,   recursive = TRUE)
if (!dir.exists(models))  dir.create(models,  recursive = TRUE)
if (!dir.exists(scripts)) dir.create(scripts, recursive = TRUE)
if (!dir.exists(data_dir)) dir.create(data_dir, recursive = TRUE)

knitr::opts_chunk$set(
  fig.path = plots
)

#################### packages ####################
# Core tidyverse for wrangling + plotting; janitor/skimr for quick structure checks; patchwork for combining plots
library(tidyverse)
library(janitor)
library(skimr)
library(knitr)
library(cowplot)
library(patchwork)
library(ggdist)
# Bayesian
library(brms);
library(loo);packageVersion("loo")
library(HDInterval);packageVersion("HDInterval")
library(cmdstanr);
library(emmeans);packageVersion("emmeans")
library(posterior);
library(bayesplot);
library(BayesFactor);
library(tidybayes);packageVersion("tidybayes")
library(bayestestR);
library(cmdstanr)

options(brms.backend = "cmdstanr") # Speed gain: ~30-50% faster than rstan
options(brms.file_refit = "on_change")  # only refit if model changes

# use all available cores for parallel computing
options(mc.cores = parallel::detectCores())
theme_set(theme_tidybayes() + panel_border());

# Keep plots/readability consistent
theme_set(theme_minimal(base_size = 14))
colorBlindBlack8  <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Figure and Table caption adapted from https://stackoverflow.com/questions/37116632/rmarkdown-html-number-figures:
outputFormat = opts_knit$get("rmarkdown.pandoc.to"); # determine the output format of the document
if( is.null(outputFormat) ) outputFormat = ""; # probably not run within knittr
capTabNo = 1; capFigNo = 1; # figure and table caption numbering, for HTML do it manually
#Function to add the Table Number
capTab = function(x)
{
  if(outputFormat == 'html'){
    x = paste0("**Table ",capTabNo,".** ",x,"")
    capTabNo <<- capTabNo + 1
  }; x
}

#Function to add the Figure Number
capFig = function(x, show_R_version=FALSE, show_package_versions=NULL, is_map=FALSE)
{
  if(outputFormat == 'html')
  {
    x <- paste0("**Figure ",capFigNo,".** ",x,"");
    if( show_R_version || (!is.null(show_package_versions) && length(show_package_versions) > 0) )
    {
      x <- paste0(x, " Figure generated using ");
      if( show_R_version ) x <- paste0(x, stringr::str_replace(R.version.string, stringr::fixed("R "), "[`R`](https://www.r-project.org/) "));
      if( !is.null(show_package_versions) && length(show_package_versions) > 0 )
      {
        x <- paste0(x, ifelse( show_R_version, " and ", " "));
        x <- paste0(x, ifelse( length(show_package_versions) > 1, "packages ", "package "));
        x <- paste0(x, paste0(vapply(show_package_versions, function(x) paste0("`",x,"`", " (version ", packageVersion(x),")"), character(1)), collapse=", "), ".");
      }
      if( is_map ) x <- paste0(x, " Maps are using public domain data from the [Natural Earth project](https://www.naturalearthdata.com/) as provided by the `R` package `maps`.");
    }
    capFigNo <<- capFigNo + 1;
  };
  x;
}

########## reproducibility ##########
set.seed(2137)
```

# Overview

**Research question.** Do speakers encode size iconically, and do listeners use those same
acoustic cues? Iconicity — a resemblance between linguistic form and meaning [@perniss_vigliocco_2014]
— has moved from a peripheral curiosity to a recognized organizing principle in language
[@cwiek_iconicity_2022; @cwiek_introduction_2025; @dingemanse_arbitrariness_2015], and size is one
of the most robust domains in which it surfaces: from the classic *mil/mal* contrast
[@sapir_study_1929] through the lexicon of English size adjectives [@winter_size_2021] and
cross-linguistic patterns [@blasi_soundmeaning_2016] to recent work on the acoustic and
articulatory bases of size–sound correspondences [@winter_size_2025].

**Main claim.** Acoustic size iconicity is primarily realized through *duration* and *pitch* (f0):
big referents are produced with longer vowels and lower f0, and listeners rely on those same cues
when categorizing referent size. The pitch prediction follows the frequency code
[@ohala_ethological_1984; @ohala_cross-language_1983], in which lower f0 indexes larger body size
across vertebrates; the duration prediction follows recent treatments of iconicity as embedded in
continuous acoustic and articulatory parameters rather than discrete phonemic features
[@winter_size_2025; @fuchs_sounds_2022; @perlman_iconicity_2018]. The same auditory analogies are
expected to guide listener categorization [@winter_size_2021; @perniss_vigliocco_2014;
@schmidt_jabeen_dragon_2024].

**Secondary claim.** Animacy enters the analysis through two channels: as a *modulator* of the iconic size effect (a `size × animacy` interaction in the production and perception models) and as an *independent main effect* on the acoustic and perceptual outcomes. The literature on iconic production is mixed, and both possibilities are kept open — interaction, main effect, both, or neither — with each tested explicitly.

**Exploratory.** Other features — F1, F2, fricative duration, fricative SPL, fricative spectral
peak, fricative centre of gravity, and the voice-quality measures CPP [@fraile_cepstral_2014;
@hillenbrand_acoustic_1994] and H1–H2 [@chai_h1h2_2022] — are examined without a priori
predictions. The pluripotential view of iconic features (cf. Akita et al., 2025; Winter, Abralin
Ao Vivo talk) motivates inspecting the broader signal rather than committing to a single channel.

This analysis lives entirely in the Bayesian framework [`brms`, @bates_fitting_2015 for the
underlying frequentist mixed-effects logic; @baayen_analyzing_2008] with the `cmdstanr`
backend. All German labels and column names from the original data files are translated to English
in the cleaning step so the script and its outputs are readable internationally.

# Decisions log

- **Primary models** include the `size × animacy` interaction so the same fits answer both the
  main and secondary claims.
- **Random effects (production)**: `(1 + size_c | speaker) + (1 | item)`. 39 speakers, 36 items,
  ~1400 vowel rows / ~2470 fricative rows — comfortable.
- **Random effects (perception)**: `(1 | listener) + (1 | speaker) + (1 | item)`. 39 × 39 × 36,
  936 observations — varying slopes added per-model only if needed.
- **Coding**: sum contrasts `-0.5 / +0.5` for `size_c` and `animacy_c` so main effects remain
  interpretable in the presence of the interaction. Continuous predictors entering perception
  models are within-speaker z-scored — the listener-side coefficient then reads as "1 SD shift in
  *this speaker's* signal".
- **Priors**: weakly informative — `normal(0, 1)` on standardized fixed effects,
  `student_t(3, 0, 2.5)` on intercepts and SDs.
- **Outcome families**:
  - durations → `lognormal()` (positive, right-skewed)
  - f0 / formants / CPP / H1H2 / SPL / peak frequency → `gaussian()`
  - binary perception (`big_judgment`, `animate_judgment`) → `bernoulli(logit)`
  - 6-way `picked_category` → `categorical()` (only as exploratory RQ3, off by default)
- **Outliers**: not trimmed by default — the lognormal family absorbs the right tail. A
  diagnostic chunk flags extreme cases for inspection.
- **Caching**: fits saved to `models/`; `brms.file_refit = "on_change"` means re-knit only
  refits when the model spec changes.
- **Acoustic extraction**: vowel features via PraatSauce [@kirby_praatsauce_2025;
  @boersma_praat_2023]; fricative features via the Al-Tamimi spectral-analyses scripts
  [@al-tamimi_jalalal-tamimipraat-spectral-analyses-fricatives-stops_2022;
  @jongman_acoustic_2000].

# 1. Read raw data

`alle.csv` is comma-separated; the others are tab-separated. `out_praatSauce.tsv` has one
malformed row and uses `--undefined--` as a missing-value sentinel.

```{r read-raw}
alle_raw <- read_csv(file.path(data_dir, "alle.csv"),
                     show_col_types = FALSE) |>
  select(-1)        # drop the unnamed row-index column

praat_raw <- read_tsv(
  file.path(data_dir, "out_praatSauce.tsv"),
  na = c("", "NA", "--undefined--"),
  show_col_types = FALSE,
  col_types = cols(
    file        = col_character(),
    label       = col_character(),
    intervaldID = col_integer(),
    .default    = col_double()       # forces f0/F1/F2/CPP/H1H2/etc. to numeric
  )
) |>
  rename(intervalID = intervaldID) |>
  filter(!is.na(intervalID))         # drops the single malformed row (line 434)

fric_raw <- read_tsv(file.path(data_dir, "spectralResultsFricatives.tsv"),
                     show_col_types = FALSE)

vdur_raw <- read_tsv(file.path(data_dir, "vowelDuration.txt"),
                     show_col_types = FALSE) |>
  rename(file = FileName, vowel = digit, vowel_dur_ms = `Duration(ms)`)
```

# 2. Translate German → English

Three layers: column names, categorical values, and (in a glossary) item words.
Stimulus item names are translated to `item_english`.

## 2.1 Column-name dictionary

```{r column-dict}
# Names of vector = original German column names; values = English replacements.
col_translations <- c(
  Gruppe              = "group",
  Bewertet            = "listener",
  Spricht             = "speaker",
  Belebtheit          = "animacy",
  Art                 = "category",
  Groesse             = "size",
  Exemplar            = "item",
  Kurz                = "item_code",
  Runde_Durchgang     = "round_trial",
  Geraten_Exemplar    = "picked_item",
  Geraten_Kurz        = "picked_item_code",
  Geraten_Belebtheit  = "picked_animacy",
  Geraten_Art         = "picked_category",
  Geraten_Groesse     = "picked_size",
  Exemplar_Korrekt    = "item_correct",
  Gruppe_Korrekt      = "group_correct",
  Belebtheit_Korrekt  = "animacy_correct",
  Art_Korrekt         = "category_correct",
  Groesse_Korrekt     = "size_correct"
)

translate_columns <- function(old_names) {
  new <- col_translations[old_names]
  ifelse(is.na(new), old_names, new)
}

alle <- alle_raw |> rename_with(translate_columns)
```

## 2.2 Value-label dictionaries

```{r value-translation}
alle <- alle |>
  mutate(
    across(c(size, picked_size),
           ~ case_when(.x == "Gross" ~ "big",
                       .x == "Klein" ~ "small",
                       TRUE ~ .x)),
    across(c(animacy, picked_animacy),
           ~ case_when(.x == "Belebt"   ~ "animate",
                       .x == "Unbelebt" ~ "inanimate",
                       TRUE ~ .x)),
    across(c(category, picked_category),
           ~ case_when(.x == "Vogel"      ~ "bird",
                       .x == "Saeugetier" ~ "mammal",
                       .x == "Reptil"     ~ "reptile",
                       .x == "Geraet"     ~ "appliance",
                       .x == "Fahrzeug"   ~ "vehicle",
                       .x == "Instrument" ~ "instrument",
                       TRUE ~ .x))
  )
```

## 2.3 Item-name glossary (German stimulus → English gloss)

```{r item-glossary}
item_glossary <- tribble(
  ~item_norm,            ~item_english,
  "buerodrucker",        "office printer",
  "drucker",             "printer",
  "industriebackofen",   "industrial oven",
  "backofen",            "oven",
  "armbanduhr",          "wristwatch",
  "wanduhr",             "wall clock",
  "kirchenglocke",       "church bell",
  "gloeckchen",          "little bell",
  "kontrabass",          "double bass",
  "violine",             "violin",
  "bodentrommel",        "floor drum",
  "handtrommel",         "hand drum",
  "grosserbagger",       "big excavator",
  "kleinerbagger",       "small excavator",
  "jumbojet",            "jumbo jet",
  "kleinflugzeug",       "small airplane",
  "suv",                 "SUV",
  "kleinwagen",          "small car",
  "elefant",             "elephant",
  "kuh",                 "cow",
  "baer",                "bear",
  "kaninchen",           "rabbit",
  "aeffchen",            "little monkey",
  "igel",                "hedgehog",
  "kondor",              "condor",
  "harpyie",             "harpy eagle",
  "storch",              "stork",
  "blaumeise",           "blue tit",
  "dompfaff",            "bullfinch",
  "spatz",               "sparrow",
  "anakonda",            "anaconda",
  "krokodil",            "crocodile",
  "komodowaran",         "komodo dragon",
  "schildkroete",        "turtle",
  "blindschleiche",      "slow worm",
  "gecko",               "gecko"
)
```

# 3. Clean and harmonize

## 3.1 Parse production filenames

Format: `group + speaker_num _ trial_num _ entity _ sex + age` (e.g. `a1_01_buerodrucker_m31`).
Two filenames lack the `_sex+age` suffix (`d1_11_anakonda`, `i1_03_kontrabass`); we recover
speaker info from same-group siblings.

```{r parse-filenames}
parse_filename <- function(x) {
  x_clean <- x |> str_remove("\\.wav$") |> str_to_lower()
  m <- str_match(x_clean,
    "^([a-z])(\\d)_(\\d{2})_(.+?)_([mw])(\\d+)$")
  tibble(
    file_key      = x_clean,
    group         = m[, 2],
    speaker_num   = as.integer(m[, 3]),
    trial_num     = as.integer(m[, 4]),
    item_norm     = m[, 5],
    sex           = m[, 6],
    age           = as.integer(m[, 7])
  )
}

# Master file table: every unique production filename across the three production sources
master_files <- bind_rows(
  praat_raw |> distinct(file)     |> rename(file_orig = file),
  fric_raw  |> distinct(fileName) |> rename(file_orig = fileName),
  vdur_raw  |> distinct(file)     |> rename(file_orig = file)
) |>
  distinct(file_orig) |>
  mutate(parsed = map(file_orig, parse_filename)) |>
  unnest(parsed)

unparsed <- master_files |> filter(is.na(group))
cat("Files that did NOT parse:\n"); print(unparsed |> select(file_orig))
```

```{r recover-unparsed}
# Recover sex/age from same-group siblings (these participants appear elsewhere).
sibling_lookup <- master_files |>
  filter(!is.na(group)) |>
  distinct(group, speaker_num, sex, age)

manual_pat <- "^([a-z])(\\d)_(\\d{2})_(.+)$"
manual <- master_files |>
  filter(is.na(group)) |>
  mutate(file_key_lc = str_to_lower(file_orig) |> str_remove("\\.wav$")) |>
  transmute(
    file_orig,
    group       = str_match(file_key_lc, manual_pat)[, 2],
    speaker_num = as.integer(str_match(file_key_lc, manual_pat)[, 3]),
    trial_num   = as.integer(str_match(file_key_lc, manual_pat)[, 4]),
    item_norm   = str_match(file_key_lc, manual_pat)[, 5],
    file_key    = file_key_lc
  ) |>
  left_join(sibling_lookup, by = c("group", "speaker_num"))

master_files <- master_files |>
  filter(!is.na(group)) |>
  bind_rows(manual) |>
  distinct(file_key, .keep_all = TRUE)   # dedupe .wav / no-extension variants

stopifnot(sum(is.na(master_files$group)) == 0)
glimpse(master_files)
```

## 3.2 Normalize item names (fix typos, harmonize spelling)

Production filenames and the perception `item` column use different conventions, plus a handful
of typos (`harypie`/`harpyie`, `industrienbackofen`/`industriebackofen`,
`schildkorete`/`schildkroete`). We normalize both sides to the same lowercase, no-spaces form.

```{r normalize-items}
normalize_item <- function(x) {
  x |>
    str_to_lower() |>
    str_trim() |>
    str_remove_all("\\s+") |>
    str_replace("^harypie$",            "harpyie") |>
    str_replace("^industrienbackofen$", "industriebackofen") |>
    str_replace("^schildkorete$",       "schildkroete")
}

master_files <- master_files |>
  mutate(item_norm = normalize_item(item_norm))

alle <- alle |>
  mutate(item_norm = normalize_item(item))

# Verify item-name correspondence after normalization
prod_items <- master_files |> distinct(item_norm) |> pull(item_norm) |> sort()
perc_items <- alle |> distinct(item_norm) |> pull(item_norm) |> sort()

cat("Items in production but not perception:\n");
print(setdiff(prod_items, perc_items))
cat("\nItems in perception but not production:\n");
print(setdiff(perc_items, prod_items))
```

## 3.3 Parse perception speaker / listener codes

```{r parse-perception-codes}
parse_speaker_code <- function(s) {
  m <- str_match(s, "^([a-z])(\\d)_([mw])_(\\d+)$")
  tibble(
    .grp     = m[, 2],
    .spknum  = as.integer(m[, 3]),
    .sex     = m[, 4],
    .age     = as.integer(m[, 5])
  )
}

alle <- alle |>
  bind_cols(parse_speaker_code(alle$speaker)  |> rename_with(~ paste0("spk", .))) |>
  bind_cols(parse_speaker_code(alle$listener) |> rename_with(~ paste0("lst", .))) |>
  separate(round_trial, into = c("round", "trial_num"),
           sep = "_", remove = FALSE, convert = TRUE) |>
  rename(speaker_num  = spk.spknum,
         speaker_sex  = spk.sex,
         speaker_age  = spk.age,
         listener_sex = lst.sex,
         listener_age = lst.age) |>
  select(-spk.grp, -lst.grp)

# Sanity: round must equal speaker_num
stopifnot(all(alle$round == alle$speaker_num))

# Binary outcomes for downstream models
alle <- alle |>
  mutate(
    big_judgment      = as.integer(picked_size      == "big"),
    animate_judgment  = as.integer(picked_animacy   == "animate"),
    portrayed_big     = as.integer(size             == "big"),
    portrayed_animate = as.integer(animacy          == "animate")
  )
```

```{r mismatch-diagnostic}
# Quality-control trace: trials where the production filename and the perception
# sheet disagree about which item was used. Per speaker verification, the
# perception sheet is authoritative — these are listed for documentation only.

trial_item_check <- master_files |>
  distinct(group, speaker_num, trial_num, item_norm) |>
  rename(item_from_filename = item_norm) |>
  inner_join(
    alle |> distinct(group, speaker_num, trial_num, item_norm) |>
      rename(item_from_perception = item_norm),
    by = c("group", "speaker_num", "trial_num")
  ) |>
  filter(item_from_filename != item_from_perception)

if (nrow(trial_item_check) > 0) {
  cat("Trial-level item-name disagreements between sources",
      "(perception sheet is authoritative; production filename ignored for item identity):\n")
  print(trial_item_check)
}
```

# 4. Build the long-format production tables

Production data are in **long format**: one row per acoustic measurement (vowel or fricative
token). Per-trial summaries used in perception models are derived from these.

Files with <3 vowels reflect real audio/production issues (per speaker verification); rows with NA vowel_dur_ms are dropped per-vowel during modelling — no global filtering needed.

```{r build-prod-long}
# ---- vowel duration: one row per vowel
vdur_long <- vdur_raw |>
  mutate(file_key = str_to_lower(file)) |>
  left_join(master_files |> select(file_key, group, speaker_num, trial_num,
                                   item_norm, sex, age),
            by = "file_key") |>
  filter(!is.na(group))

# ---- praatSauce: one row per vowel, with f0 / F1 / F2 / CPP / H1-H2.
# H1-H2 = A1c - A2c per the readme.
praat_long <- praat_raw |>
  mutate(file_key = str_to_lower(file) |> str_remove("\\.wav$"),
         H1H2     = A1c - A2c) |>
  rename(vowel = label) |>
  left_join(master_files |> select(file_key, group, speaker_num, trial_num,
                                   item_norm, sex, age),
            by = "file_key") |>
  filter(!is.na(group))

# Join vowel duration onto praat (per file × vowel)
vowel_long <- praat_long |>
  left_join(vdur_long |> select(file_key, vowel, vowel_dur_ms),
            by = c("file_key", "vowel"))

# ---- Fricatives: one row per fricative token
fric_long <- fric_raw |>
  mutate(file_key = str_to_lower(fileName) |> str_remove("\\.wav$")) |>
  rename(fric_dur_ms  = durationMS,
         fric_SPL_dB  = levelMWholedB,    # whole-segment, mid-band intensity
         fric_peak_Hz = peakFreqHzMid,
         fric_CoG_Hz  = cogMid) |>
  filter(phoneme %in% c("C", "f", "z", "S")) |>   # drop one anomalous "´z" record
  left_join(master_files |> select(file_key, group, speaker_num, trial_num,
                                   item_norm, sex, age),
            by = "file_key") |>
  filter(!is.na(group)) |>
  select(file_key, phoneme, fric_dur_ms, fric_SPL_dB, fric_peak_Hz, fric_CoG_Hz,
         group, speaker_num, trial_num, item_norm, sex, age)
```

## 4.1 Attach experimental factors and item glosses to production rows

```{r attach-factors}
trial_factors <- alle |>
  distinct(group, speaker_num, trial_num, item_norm,
           animacy, category, size, item_code, item) |>
  rename(item_label = item)

speaker_id <- master_files |>
  distinct(group, speaker_num, sex, age) |>
  mutate(speaker = paste0(group, speaker_num, "_", sex, "_", age))

vowel_long <- vowel_long |>
  select(-item_norm) |>                       # filename's label is unreliable; perception wins
  left_join(trial_factors, by = c("group", "speaker_num", "trial_num")) |>
  left_join(speaker_id,    by = c("group", "speaker_num", "sex", "age")) |>
  left_join(item_glossary, by = "item_norm")

fric_long <- fric_long |>
  select(-item_norm) |>
  left_join(trial_factors, by = c("group", "speaker_num", "trial_num")) |>
  left_join(speaker_id,    by = c("group", "speaker_num", "sex", "age")) |>
  left_join(item_glossary, by = "item_norm")

cat("vowel_long rows missing factors:", sum(is.na(vowel_long$size)), "\n")
cat("fric_long  rows missing factors:", sum(is.na(fric_long$size)),  "\n")
```

## 4.2 Build the perception dataframe with speaker-level acoustic predictors

For each (group, speaker, trial), summarize the speaker's acoustic output for that production —
mean across vowels for vocalic features, mean across fricatives for fricative features. Predictors
are then within-speaker z-scored so the perception coefficient reads as "what does +1 SD of *this
speaker's* signal do to listener choice?".

```{r build-perception}
# Per-trial vowel summaries
trial_vowel <- vowel_long |>
  group_by(group, speaker_num, trial_num) |>
  summarise(
    vowel_dur_ms_mean = mean(vowel_dur_ms, na.rm = TRUE),
    f0_mean           = mean(f0,           na.rm = TRUE),
    F1_mean           = mean(F1,           na.rm = TRUE),
    F2_mean           = mean(F2,           na.rm = TRUE),
    CPP_mean          = mean(CPP,          na.rm = TRUE),
    H1H2_mean         = mean(H1H2,         na.rm = TRUE),
    .groups = "drop"
  )

# Per-trial fricative summaries
trial_fric <- fric_long |>
  group_by(group, speaker_num, trial_num) |>
  summarise(
    fric_dur_ms_mean  = mean(fric_dur_ms,  na.rm = TRUE),
    fric_SPL_dB_mean  = mean(fric_SPL_dB,  na.rm = TRUE),
    fric_peak_Hz_mean = mean(fric_peak_Hz, na.rm = TRUE),
    fric_CoG_Hz_mean  = mean(fric_CoG_Hz,  na.rm = TRUE),
    .groups = "drop"
  )

zscale <- function(x) as.numeric(scale(x))

perc <- alle |>
  left_join(trial_vowel,   by = c("group", "speaker_num", "trial_num")) |>
  left_join(trial_fric,    by = c("group", "speaker_num", "trial_num")) |>
  left_join(item_glossary, by = "item_norm") |>
  group_by(speaker) |>
  mutate(across(c(vowel_dur_ms_mean, f0_mean, F1_mean, F2_mean, CPP_mean, H1H2_mean,
                  fric_dur_ms_mean, fric_SPL_dB_mean, fric_peak_Hz_mean, fric_CoG_Hz_mean),
                zscale, .names = "{.col}_z")) |>
  ungroup()

cat("Perception rows without production audio:",
    sum(is.na(perc$vowel_dur_ms_mean)), "of", nrow(perc), "\n")
```

## 4.3 Sum-coding for size and animacy

```{r factor-coding}
# Helper: 2-level sum-coded contrast (-0.5 / +0.5)
sum_code_2 <- function(x, levels) {
  x <- factor(x, levels = levels)
  as.numeric(x) - 1.5      # level 1 -> -0.5, level 2 -> +0.5
}

# Production-side
vowel_long <- vowel_long |>
  mutate(
    size_c    = sum_code_2(size,    c("small",     "big")),       # +0.5 = big
    animacy_c = sum_code_2(animacy, c("inanimate", "animate"))    # +0.5 = animate
  )

fric_long <- fric_long |>
  mutate(
    size_c    = sum_code_2(size,    c("small",     "big")),
    animacy_c = sum_code_2(animacy, c("inanimate", "animate"))
  )

# Perception-side
perc <- perc |>
  mutate(
    size_c    = sum_code_2(size,    c("small",     "big")),
    animacy_c = sum_code_2(animacy, c("inanimate", "animate"))
  )
```

# 5. Exploratory data analysis

## 5.1 Sample sizes

```{r eda-counts}
cat("Vowel rows:        ", nrow(vowel_long), "\n")
cat("Fricative rows:    ", nrow(fric_long),  "\n")
cat("Perception rows:   ", nrow(perc),       "\n")
cat("Speakers:          ", n_distinct(perc$speaker),   "\n")
cat("Listeners:         ", n_distinct(perc$listener),  "\n")
cat("Items:             ", n_distinct(perc$item_english), "\n")

perc |> tabyl(animacy, size, category) |> print()
```

## 5.2 Distributions of acoustic features

```{r eda-distributions, fig.height=6, fig.cap=capFig("Vocalic features by size and animacy. Median lines inside violins.")}
acoustic_long <- vowel_long |>
  select(speaker, item_english, vowel, size, animacy,
         vowel_dur_ms, f0, F1, F2, CPP, H1H2) |>
  pivot_longer(c(vowel_dur_ms, f0, F1, F2, CPP, H1H2),
               names_to = "feature", values_to = "value")

ggplot(acoustic_long, aes(x = size, y = value, fill = animacy)) +
  geom_violin(draw_quantiles = 0.5, alpha = 0.7) +
  facet_wrap(~ feature, scales = "free_y") +
  scale_fill_manual(values = colorBlindBlack8[c(3, 4)]) +
  labs(title = "Vocalic features by size and animacy",
       x = NULL, y = NULL)
```

```{r eda-fricatives, fig.height=4, fig.cap=capFig("Fricative features by size and animacy.")}
fric_long |>
  pivot_longer(c(fric_dur_ms, fric_SPL_dB, fric_peak_Hz, fric_CoG_Hz),
               names_to = "feature", values_to = "value") |>
  ggplot(aes(x = size, y = value, fill = animacy)) +
  geom_violin(draw_quantiles = 0.5, alpha = 0.7) +
  facet_wrap(~ feature, scales = "free_y") +
  scale_fill_manual(values = colorBlindBlack8[c(3, 4)]) +
  labs(title = "Fricative features by size and animacy",
       x = NULL, y = NULL)
```

## 5.3 Outlier flags (vowel duration)

```{r outlier-flags}
vowel_long |>
  group_by(vowel) |>
  mutate(z = scale(log(vowel_dur_ms))[, 1]) |>
  ungroup() |>
  filter(abs(z) > 3) |>
  select(file_key, vowel, vowel_dur_ms, z, size, animacy) |>
  arrange(desc(abs(z)))
```

Extreme productions are retained: speaker verification confirmed these are genuine iconic productions (not measurement errors). Trimming would discard signal. The lognormal family on duration accommodates the right tail; PPCs are the diagnostic that matters.

## 5.4 Perception accuracy snapshot

```{r eda-accuracy}
perc |>
  summarise(
    size_acc      = mean(size_correct),
    animacy_acc   = mean(animacy_correct),
    category_acc  = mean(category_correct),
    item_acc      = mean(item_correct)
  )
```

```{r eda-perc-by-cell, fig.height=4, fig.cap=capFig("Proportion of 'big' judgments by portrayed size and animacy of the referent.")}
perc |>
  group_by(animacy, size) |>
  summarise(prop_big_judgment = mean(big_judgment), .groups = "drop") |>
  ggplot(aes(x = size, y = prop_big_judgment, fill = animacy)) +
  geom_col(position = position_dodge()) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  scale_fill_manual(values = colorBlindBlack8[c(3, 4)]) +
  labs(y = "P(\"big\" judgment)",
       title = "Listener \"big\" judgment by portrayed size and animacy")
```

# 6. Primary analyses — duration and pitch

The primary models test the **main claim** (size iconicity through duration and f0
[@ohala_ethological_1984; @winter_size_2025; @fuchs_sounds_2022]) via the main effect of
`size_c`, and the **secondary claim** — that animacy enters the acoustic and perceptual picture — via both the `size_c × animacy_c` interaction (does animacy modulate the size effect?) and the `animacy_c` main effect (does animacy shift the outcome independent of size?). Including both in the same model lets a single posterior speak to both claims and avoids the multiple-testing pitfalls of fitting them separately.

## 6.1 A reusable fitter

```{r helper-fit}
default_priors <- function(family = "gaussian") {
  switch(family,
    gaussian  = c(prior(normal(0, 100),        class = "b"),
                  prior(student_t(3, 0, 100),  class = "Intercept"),
                  prior(student_t(3, 0, 100),  class = "sd"),
                  prior(student_t(3, 0, 100),  class = "sigma")),
    lognormal = c(prior(normal(0, 1.5),        class = "b"),
                  prior(student_t(3, 5, 5),    class = "Intercept"),
                  prior(student_t(3, 0, 2),    class = "sd"),
                  prior(student_t(3, 0, 2),    class = "sigma")),
    bernoulli = c(prior(normal(0, 5),          class = "b"),
                  prior(student_t(3, 0, 10),   class = "Intercept"),
                  prior(student_t(3, 0, 5),    class = "sd"))
  )
}

fit_brm <- function(formula, data, family = gaussian(), file = NULL,
                    priors = NULL, ...) {
  fam_name <- family$family
  if (is.null(priors)) priors <- default_priors(fam_name)
  brm(formula = formula, data = data, family = family,
      prior = priors,
      chains = 4, iter = 8000, warmup = 4000, cores = 4,
      control = list(adapt_delta = 0.95, max_treedepth = 12),
      seed   = 2137,
      file   = file,
      refresh = 100,
      ...)
}
```

## 6.2 Production: vowel duration

```{r m-vowel-dur, results='hide'}
m_vdur <- fit_brm(
  bf(vowel_dur_ms ~ size_c * animacy_c + vowel +
       (1 + size_c | speaker) + (1 | item_english)),
  data   = vowel_long |> filter(!is.na(vowel_dur_ms)),
  family = lognormal(),
  file   = paste0(models, "m_vdur")
)
```

```{r m-vowel-dur-ppc, fig.cap=capFig("Posterior predictive check for vowel duration (lognormal).")}
summary(m_vdur)
pp_check(m_vdur, ndraws = 100) + ggtitle("PPC — vowel duration")
```

## 6.3 Production: f0 (pitch)

f0 is modeled with `gaussian()` plus a `sex` covariate (male/female speakers differ by ~80–100 Hz
on average; without `sex` the random intercept absorbs it but you lose interpretability of the
fixed-effect intercept). The prediction — lower f0 for big referents — is the central frequency-code
expectation [@ohala_ethological_1984; @ohala_cross-language_1983].

```{r m-f0, results='hide'}
m_f0 <- fit_brm(
  bf(f0 ~ size_c * animacy_c + sex + vowel +
       (1 + size_c | speaker) + (1 | item_english)),
  data   = vowel_long |> filter(!is.na(f0)),
  family = lognormal(),
  file   = paste0(models, "m_f0")
)
```

```{r m-f0-ppc, fig.cap=capFig("Posterior predictive check for f0 (lognormal, with sex covariate).")}
summary(m_f0)
pp_check(m_f0, ndraws = 100) + ggtitle("PPC — f0")
```

## 6.4 Perception: vowel duration → "big" judgment

```{r m-perc-dur, results='hide'}
m_perc_dur <- fit_brm(
  bf(big_judgment ~ vowel_dur_ms_mean_z * animacy_c +
       (1 | listener) + (1 | speaker) + (1 | item_english)),
  data   = perc |> filter(!is.na(vowel_dur_ms_mean_z)),
  family = bernoulli(),
  file   = paste0(models, "m_perc_dur")
)
```

```{r m-perc-dur-ppc, fig.cap=capFig("Posterior predictive check for perception of duration (Bernoulli).")}
summary(m_perc_dur)
pp_check(m_perc_dur, ndraws = 100) + ggtitle("PPC — duration perception")
```

## 6.5 Perception: f0 → "big" judgment

```{r m-perc-f0, results='hide'}
m_perc_f0 <- fit_brm(
  bf(big_judgment ~ f0_mean_z * animacy_c +
       (1 | listener) + (1 | speaker) + (1 | item_english)),
  data   = perc |> filter(!is.na(f0_mean_z)),
  family = bernoulli(),
  file   = paste0(models, "m_perc_f0")
)
```

```{r m-perc-f0-ppc, fig.cap=capFig("Posterior predictive check for perception of duration (Bernoulli).")}
summary(m_perc_f0)
pp_check(m_perc_f0, ndraws = 100) + ggtitle("PPC — f0 perception")
```

## 6.6 Reporting helpers

```{r report-helper}
report_model <- function(model, label = "model", terms_of_interest = NULL) {
  cat("\n========== ", label, " ==========\n")

  ds <- describe_posterior(model, ci = 0.95, ci_method = "hdi",
                           test = c("pd"), centrality = "median",
                           effects = "fixed")
  print(ds)

  print(pp_check(model, ndraws = 100) + ggtitle(paste("PPC —", label)))

  draws <- as_draws_df(model) |>
    select(starts_with("b_"), -matches("^b_Intercept$")) |>
    pivot_longer(everything(), names_to = "term", values_to = "value") |>
    mutate(term = str_remove(term, "^b_"))

  if (!is.null(terms_of_interest))
    draws <- draws |> filter(term %in% terms_of_interest)

  p <- ggplot(draws, aes(x = value, y = term)) +
    stat_halfeye(.width = c(0.66, 0.95)) +
    geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
    labs(title = paste("Posterior —", label), x = "estimate", y = NULL)
  print(p)

  invisible(ds)
}

report_effect <- function(model, coef_name,
                          direction = c("auto", "positive", "negative"),
                          label = NULL) {
  direction   <- match.arg(direction)
  family_name <- family(model)$family
  if (is.null(label)) label <- paste0(deparse(substitute(model)), " :: ", coef_name)

  # ---- Posterior draws of the coefficient on the model's link scale
  beta <- as_draws_df(model)[[paste0("b_", coef_name)]]
  if (is.null(beta)) stop("Coefficient '", coef_name, "' not in model.")

  link_hdi <- median_hdi(beta, .width = 0.95)

  # ---- Decide direction for the one-sided hypothesis test
  if (direction == "auto") {
    direction <- if (median(beta) > 0) "positive" else "negative"
  }
  hyp_str <- paste(coef_name, if (direction == "positive") "> 0" else "< 0")
  hyp     <- hypothesis(model, hyp_str)$hypothesis

  # ---- Print
  cat("\n--- ", label, "  [", family_name, "] ---\n", sep = "")
  cat(sprintf("Link-scale coefficient:  %.4f   [95%% HDI: %.4f, %.4f]\n",
              link_hdi$y, link_hdi$ymin, link_hdi$ymax))

  if (family_name == "lognormal") {
    # With sum-coded predictor (-0.5/+0.5), beta is exactly log(y_big / y_small).
    # So exp(beta) is the ratio big/small, and (exp(beta) - 1) * 100 is the % change.
    pct      <- (exp(beta) - 1) * 100
    pct_hdi  <- median_hdi(pct, .width = 0.95)
    ratio_hdi <- median_hdi(exp(beta), .width = 0.95)
    cat(sprintf("Multiplicative effect:   %.3f×  [95%% HDI: %.3f×, %.3f×]\n",
                ratio_hdi$y, ratio_hdi$ymin, ratio_hdi$ymax))
    cat(sprintf("As %% change (big - small): %.2f%% [95%% HDI: %.2f%%, %.2f%%]\n",
                pct_hdi$y, pct_hdi$ymin, pct_hdi$ymax))
  } else if (family_name == "bernoulli") {
    or     <- exp(beta)
    or_hdi <- median_hdi(or, .width = 0.95)
    cat(sprintf("Odds ratio:              %.3f   [95%% HDI: %.3f, %.3f]\n",
                or_hdi$y, or_hdi$ymin, or_hdi$ymax))
  }
  # For gaussian the link IS the response scale, so nothing to back-transform.

  cat(sprintf("P(%s | data)  =  %.4f\n", hyp_str, hyp$Post.Prob))

  invisible(list(link = link_hdi, hypothesis = hyp))
}
```


```{r}
# Main claim: size effects
report_effect(m_vdur,     "size_c",                direction = "positive",
              label = "H1.1pro vowel duration ~ size")
report_effect(m_f0,       "size_c",                direction = "negative",
              label = "H2.1pro f0 ~ size")
report_effect(m_perc_dur, "vowel_dur_ms_mean_z",   direction = "positive",
              label = "H1.1per duration → big judgment")
report_effect(m_perc_f0,  "f0_mean_z",             direction = "negative",
              label = "H2.1per f0 → big judgment")

# Secondary claim: animacy effects (modulation and main effects)
report_effect(m_vdur, "size_c:animacy_c", direction = "auto",
              label = "Animacy modulation of vowel duration")
report_effect(m_f0,   "size_c:animacy_c", direction = "auto",
              label = "Animacy modulation of f0")

report_effect(m_vdur,     "animacy_c", direction = "auto",
              label = "Animacy main effect on vowel duration")
report_effect(m_f0,       "animacy_c", direction = "auto",
              label = "Animacy main effect on f0")
report_effect(m_perc_dur, "animacy_c", direction = "auto",
              label = "Animacy main effect on 'big' judgment (duration model)")
report_effect(m_perc_f0,  "animacy_c", direction = "auto",
              label = "Animacy main effect on 'big' judgment (f0 model)")
```


```{r report-primary}
walk2(
  list(m_vdur, m_f0, m_perc_dur, m_perc_f0),
  c("Production: vowel duration",
    "Production: f0",
    "Perception: duration → big judgment",
    "Perception: f0 → big judgment"),
  ~ report_model(.x, label = .y)
)
```

### Summary of primary findings

**Production — size iconicity (main claim, supported).**
Speakers produced big referents with vowels **~18% longer** than small referents
(β = 0.16 on the log scale, 95% HDI [0.02, 0.30]; P(β > 0) = 0.99) and with
**~17% lower f0** (β = −0.18, 95% HDI [−0.28, −0.08]; P(β < 0) = 1.00).
Both production hypotheses (H1.1pro and H2.1pro) are supported.

**Perception — only the f0 loop closes (mixed support).**
Listeners used **f0 as a cue to size**: lower f0 increased the odds of a "big"
judgment by a factor of ~0.60 per +1 SD of speaker f0 (β = −0.52, 95% HDI
[−0.77, −0.25]; OR = 0.60; P(β < 0) ≈ 1.00). H2.1per is supported.
Listeners did **not** show a comparable effect for vowel duration
(β = 0.05, 95% HDI [−0.14, 0.25]; OR = 1.05; P(β > 0) = 0.70). H1.1per is not
supported. The frequency-code channel is bidirectional in this dataset (produced
and perceived); the duration channel is asymmetric — speakers produce iconic
lengthening for big referents, but listeners do not appear to use vowel
duration as a cue to size.

**Animacy modulation (secondary claim, not supported).**
The size × animacy interaction was close to zero on both acoustic dimensions
(vowel duration: β = 0.04, 95% HDI [−0.22, 0.30], P = 0.61; f0: β = −0.11,
95% HDI [−0.31, 0.08], P = 0.87). The size effect did not differ measurably
between animate and inanimate referents — iconic size encoding operates the
same way across both categories.

**Animacy main effect (unanticipated finding, supported in production).**
Independent of size, animate referents were produced with vowels **~14% shorter**
than inanimate referents (β = −0.15, 95% HDI [−0.28, −0.02]; P(β < 0) = 0.99) and
with **~11% higher f0** (β = 0.10, 95% HDI [0.00, 0.20]; P(β > 0) = 0.98).
On the perception side both animacy main effects pointed in the same direction
(animate referents less likely to be judged "big") but their credible intervals
crossed zero (duration model: β = −0.92, 95% HDI [−2.18, 0.36], P = 0.93;
f0 model: β = −0.71, 95% HDI [−1.78, 0.39], P = 0.91).

**Reading the main-effect-without-interaction pattern.** That speakers shift
their acoustic baseline by animacy *without* changing how strongly they encode
size means two things happen in parallel: (i) the size mapping is robust across
the animacy distinction, and (ii) there is a separate, additive register-like
shift between speaking for animals and speaking for objects. Whether this
shift reflects animacy as a feature or systematic differences between the
animal and object item sets in this stimulus battery cannot be separated with
the present design.

## 6.7 Main claim — size iconicity main effects

A single forest plot of the size posterior across the four primary models — the headline graphic
for the main claim. Positive values indicate the size-iconic direction for duration (longer for
big); for f0 the prediction is **negative** (lower for big), so a posterior concentrated below
zero supports the frequency-code prediction [@ohala_ethological_1984; @winter_size_2025].

```{r forest-main, fig.height=3.5, fig.cap=capFig("Headline plot — size iconicity main effects across the four primary models. Filled intervals: 66% and 95% HDI; dashed line at zero.")}
extract_term <- function(model, term, name) {
  as_draws_df(model) |>
    transmute(model = name, value = .data[[paste0("b_", term)]])
}

main_effects <- bind_rows(
  extract_term(m_vdur,     "size_c",                "vowel duration (production, log-ms)"),
  extract_term(m_f0,       "size_c",                "f0 (production, Hz)"),
  extract_term(m_perc_dur, "vowel_dur_ms_mean_z",   "duration → big (perception, log-odds)"),
  extract_term(m_perc_f0,  "f0_mean_z",             "f0 → big (perception, log-odds)")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(main_effects, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "Effect estimate (sum-coded big vs. small / per +1 SD)",
       y = NULL,
       title = "Main claim: size iconicity in production and perception",
       subtitle = "Predicted direction: positive for duration; negative for f0")
```

## 6.8 Secondary claim — animacy effects

The model contains two animacy-relevant terms: the **interaction** `size_c × animacy_c`
(does the size effect differ between animate and inanimate referents?) and the
**main effect** of `animacy_c` (does animacy shift the outcome independent of size?).
A finding can land on either, both, or neither. Reporting both keeps us honest about
what the data say.

```{r forest-interaction, fig.height=3.5, fig.cap=capFig("Size × animacy interaction across the four primary models.")}
interactions_df <- bind_rows(
  extract_term(m_vdur,     "size_c:animacy_c",                  "vowel duration (production)"),
  extract_term(m_f0,       "size_c:animacy_c",                  "f0 (production)"),
  extract_term(m_perc_dur, "vowel_dur_ms_mean_z:animacy_c",     "duration (perception)"),
  extract_term(m_perc_f0,  "f0_mean_z:animacy_c",               "f0 (perception)")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(interactions_df, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "Size × Animacy interaction estimate",
       y = NULL,
       title = "Does animacy modulate the size effect?",
       subtitle = "Nonzero = size effect differs between animate and inanimate referents")
```

```{r forest-animacy-main, fig.height=3.5, fig.cap=capFig("Animacy main effect across the four primary models.")}
animacy_main_df <- bind_rows(
  extract_term(m_vdur,     "animacy_c", "vowel duration (production)"),
  extract_term(m_f0,       "animacy_c", "f0 (production)"),
  extract_term(m_perc_dur, "animacy_c", "duration model (perception)"),
  extract_term(m_perc_f0,  "animacy_c", "f0 model (perception)")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(animacy_main_df, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "Animacy main effect estimate",
       y = NULL,
       title = "Does animacy shift the outcome independent of size?",
       subtitle = "Nonzero = animate and inanimate referents have different baselines")
```

```{r parallel-lines-helper}
# Predicted outcome at the four corners of size × animacy.
# Random effects are integrated out (re_formula = NA): we get population-level
# predictions, which is what makes the parallel-vs-tilted geometry clean.
plot_size_by_animacy <- function(model, y_label, title_text,
                                 vowel_ref = "a:", phoneme_ref = "C", sex_ref = "m") {
  # Use the model's actual data columns to decide which covariates to supply.
  # all.vars(formula(model)) doesn't reliably traverse a brmsformula tree.
  model_vars <- names(model$data)

  newd <- expand_grid(
    size_c    = c(-0.5, 0.5),
    animacy_c = c(-0.5, 0.5)
  )
  if ("vowel"   %in% model_vars) newd$vowel   <- vowel_ref
  if ("phoneme" %in% model_vars) newd$phoneme <- phoneme_ref
  if ("sex"     %in% model_vars) newd$sex     <- sex_ref

  newd |>
    add_epred_draws(model, re_formula = NA) |>
    mutate(
      size    = factor(if_else(size_c < 0, "small", "big"), levels = c("small", "big")),
      animacy = if_else(animacy_c < 0, "inanimate", "animate")
    ) |>
    ggplot(aes(x = size, y = .epred, colour = animacy, group = animacy)) +
    stat_pointinterval(.width = c(0.66, 0.95),
                       position = position_dodge(width = 0.2)) +
    stat_summary(fun = median, geom = "line",
                 position = position_dodge(width = 0.2),
                 linewidth = 0.8) +
    scale_colour_manual(values = colorBlindBlack8[c(3, 4)]) +
    labs(x = NULL, y = y_label, colour = "Animacy", title = title_text)
}
```

```{r size-by-animacy-plots, fig.width=10, fig.height=4.5, fig.cap=capFig("Population-level predicted outcomes at the four cells of size × animacy. Parallel lines indicate an animacy main effect without modulation; tilted (non-parallel) lines would indicate that the size effect itself differs between animate and inanimate referents.")}
p1 <- plot_size_by_animacy(m_vdur, "Vowel duration (ms)", "Vowel duration")
p2 <- plot_size_by_animacy(m_f0,   "f0 (Hz)",             "f0")

(p1 + p2) + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

Look at the *vowel duration* panel: green (inanimate) sits above blue (animate), both climb at essentially the same rate from small → big. That's the textbook parallel-lines pattern — animacy shifts the baseline (the ~14% gap), size effect is the same for both (the ~17% lift). This is the `m_vdur` model: significant animacy main effect, interaction at zero. 

The f0 panel tells a slightly different story. The lines aren't quite parallel — blue (animate) starts higher at `small` (~**148** Hz) and converges with green (inanimate) at `big` (~**115** Hz). That mild convergence is exactly the soft hint of a non-zero interaction we saw in the numbers (β = −0.11, 95% **HDI** [−0.31, 0.08]; P(β < 0) = 0.87). The CrI brushes zero so it's not a claim, but the geometry makes the hint visible: animate referents seem to do more iconic "lowering" than inanimate ones do.

# 7. Exploratory analyses

These models reuse the primary structure with swapped outcomes/predictors. They are deliberately exploratory: examining the broader acoustic signal for where iconicity may leak into less-prototyped channels, without a priori predictions and without claim about robustness pending replication.

## 7.1 A pair of templates

```{r exploratory-template}
# Production: continuous vocalic outcome
fit_prod_vocalic <- function(outcome, family = gaussian(), data = vowel_long) {
  formula <- bf(as.formula(paste(
    outcome, "~ size_c * animacy_c + sex + vowel +",
    "(1 + size_c | speaker) + (1 | item_english)"
  )))
  d <- data |> filter(!is.na(.data[[outcome]]))
  fit_brm(formula, data = d, family = family,
          file = paste0(models, "m_prod_", outcome))
}

# Production: continuous fricative outcome
fit_prod_fric <- function(outcome, family = gaussian(), data = fric_long) {
  formula <- bf(as.formula(paste(
    outcome, "~ size_c * animacy_c + phoneme +",
    "(1 + size_c | speaker) + (1 | item_english)"
  )))
  d <- data |> filter(!is.na(.data[[outcome]]))
  fit_brm(formula, data = d, family = family,
          file = paste0(models, "m_prod_", outcome))
}

# Perception: binary outcome on a single z-scored predictor
fit_perc_bin <- function(outcome, predictor_z, data = perc) {
  formula <- bf(as.formula(paste(
    outcome, "~", predictor_z, "* animacy_c +",
    "(1 | listener) + (1 | speaker) + (1 | item_english)"
  )))
  d <- data |> filter(!is.na(.data[[predictor_z]]))
  fit_brm(formula, data = d, family = bernoulli(),
          file = paste0(models, "m_perc_", outcome, "_", predictor_z))
}
```

## 7.2 Other vocalic features (F1, F2, CPP, H1–H2)

```{r exploratory-vocalic, results='hide'}
m_F1   <- fit_prod_vocalic("F1",   family = lognormal())
m_F2   <- fit_prod_vocalic("F2",   family = lognormal())
m_CPP  <- fit_prod_vocalic("CPP")          # gaussian (default)
m_H1H2 <- fit_prod_vocalic("H1H2")         # gaussian (default)
```

F1 has been examined as a size-iconic cue in production by Ćwiek & Fuchs (2025), who reported no systematic mapping from referent size to F1 (or lip opening) in a directed-speech paradigm with inanimate referents.

## 7.3 Fricative features (production)

Fricative spectral cues are routinely linked to perceived size and to the geometry of the vocal
tract [@jongman_acoustic_2000; @al-tamimi_jalalal-tamimipraat-spectral-analyses-fricatives-stops_2022].

```{r exploratory-fric, results='hide'}
m_fdur  <- fit_prod_fric("fric_dur_ms",  family = lognormal())
m_fSPL  <- fit_prod_fric("fric_SPL_dB")              # gaussian
m_fpeak <- fit_prod_fric("fric_peak_Hz")             # gaussian (log overshoots)
m_fCoG  <- fit_prod_fric("fric_CoG_Hz")              # gaussian (log overshoots)
```

## 7.4 Perception of exploratory features

```{r exploratory-perc, results='hide'}
m_perc_F1     <- fit_perc_bin("big_judgment", "F1_mean_z")
m_perc_F2     <- fit_perc_bin("big_judgment", "F2_mean_z")
m_perc_fdur   <- fit_perc_bin("big_judgment", "fric_dur_ms_mean_z")
m_perc_fSPL   <- fit_perc_bin("big_judgment", "fric_SPL_dB_mean_z")
m_perc_fpeak  <- fit_perc_bin("big_judgment", "fric_peak_Hz_mean_z")
```

## 7.5 RQ4 — does animacy *judgment* depend on acoustic cues?

```{r rq4, results='hide'}
m_anim_dur <- fit_perc_bin("animate_judgment", "vowel_dur_ms_mean_z")
m_anim_f0  <- fit_perc_bin("animate_judgment", "f0_mean_z")
```

## 7.6 RQ3 — multinomial subcategory judgment

Categorical (6-level) outcome models are slow. Off by default; flip `eval=TRUE` to run.

```{r rq3, eval=FALSE, results='hide'}
m_cat_dur <- fit_brm(
  bf(picked_category ~ vowel_dur_ms_mean_z * animacy_c +
       (1 | listener) + (1 | speaker) + (1 | item_english)),
  data   = perc |> filter(!is.na(vowel_dur_ms_mean_z)) |>
                   mutate(picked_category = factor(picked_category)),
  family = categorical(),
  file   = paste0(models, "m_cat_dur")
)
```

# 8. Cross-feature comparison

## 8.1 Production: where does size leak into the signal?

```{r forest-prod, fig.height=5, fig.cap=capFig("Size effect across all production models. Top two rows = primary claim (duration and f0); below = exploratory features.")}
forest_prod <- bind_rows(
  extract_term(m_vdur,  "size_c", "vowel duration (log)"),
  extract_term(m_f0,    "size_c", "f0 (Hz)"),
  extract_term(m_F1,    "size_c", "F1 (Hz)"),
  extract_term(m_F2,    "size_c", "F2 (Hz)"),
  extract_term(m_CPP,   "size_c", "CPP (dB)"),
  extract_term(m_H1H2,  "size_c", "H1–H2"),
  extract_term(m_fdur,  "size_c", "fricative duration (log)"),
  extract_term(m_fSPL,  "size_c", "fricative SPL (dB)"),
  extract_term(m_fpeak, "size_c", "fricative peak (Hz)"),
  extract_term(m_fCoG,  "size_c", "fricative CoG (Hz)")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(forest_prod, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "Effect of big vs. small (sum-coded)",
       y = NULL,
       title = "Production: size effect across all acoustic features",
       subtitle = "Top two rows = primary claim; below = exploratory")
```

## 8.2 Perception: which cues drive "big" judgments?

```{r forest-perc, fig.height=4, fig.cap=capFig("Perception: log-odds shift in 'big' judgment per +1 SD of each acoustic feature. Top two rows = primary claim; below = exploratory.")}
forest_perc <- bind_rows(
  extract_term(m_perc_dur,   "vowel_dur_ms_mean_z", "vowel duration"),
  extract_term(m_perc_f0,    "f0_mean_z",           "f0"),
  extract_term(m_perc_F1,    "F1_mean_z",           "F1"),
  extract_term(m_perc_F2,    "F2_mean_z",           "F2"),
  extract_term(m_perc_fdur,  "fric_dur_ms_mean_z",  "fricative duration"),
  extract_term(m_perc_fSPL,  "fric_SPL_dB_mean_z",  "fricative SPL"),
  extract_term(m_perc_fpeak, "fric_peak_Hz_mean_z", "fricative peak")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(forest_perc, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "log-odds shift of \"big\" judgment per +1 SD acoustic feature",
       y = NULL,
       title = "Perception: which cues drive \"big\" judgments?",
       subtitle = "Top two rows = primary claim; below = exploratory")
```

# 9. References {-}

<div id="refs"></div>

# 10. Session info

```{r session}
sessionInfo()
```
