---
title: "Size symbolism in production and perception"
subtitle: "A Bayesian analysis of duration, f0, and animacy"
author: anonymized for peer review
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    df_print: paged
bibliography: MAIconicity.bib
link-citations: true
editor_options:
  chunk_output_type: console
---

```{r source setup, echo = TRUE, message=FALSE, warning = FALSE}
########## folders ##########
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
parentfolder  <- dirname(getwd())
models        <- paste0(parentfolder, '/models/')
plots         <- paste0(parentfolder, '/plots/')
scripts       <- paste0(parentfolder, '/scripts/')
data_dir      <- paste0(parentfolder, '/data/')   # adjust if data lives elsewhere

########## folder creation (security: only if missing) ##########
if (!dir.exists(plots))   dir.create(plots,   recursive = TRUE)
if (!dir.exists(models))  dir.create(models,  recursive = TRUE)
if (!dir.exists(scripts)) dir.create(scripts, recursive = TRUE)
if (!dir.exists(data_dir)) dir.create(data_dir, recursive = TRUE)

knitr::opts_chunk$set(
  fig.path = plots
)

#################### packages ####################
# Core tidyverse for wrangling + plotting; janitor/skimr for quick structure checks; patchwork for combining plots
library(tidyverse)
library(janitor)
library(skimr)
library(knitr)
library(cowplot)
library(patchwork)
library(ggdist)
# Bayesian
library(brms);
library(loo);packageVersion("loo")
library(HDInterval);packageVersion("HDInterval")
library(cmdstanr);
library(emmeans);packageVersion("emmeans")
library(posterior);
library(bayesplot);
library(BayesFactor);
library(tidybayes);packageVersion("tidybayes")
library(bayestestR);
library(cmdstanr)

options(brms.backend = "cmdstanr") # Speed gain: ~30-50% faster than rstan
options(brms.file_refit = "on_change")  # only refit if model changes

# use all available cores for parallel computing
options(mc.cores = parallel::detectCores())
theme_set(theme_tidybayes() + panel_border());

# Keep plots/readability consistent
theme_set(theme_minimal(base_size = 14))
colorBlindBlack8  <- c("#000000", "#E69F00", "#56B4E9", "#009E73",
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Figure and Table caption adapted from https://stackoverflow.com/questions/37116632/rmarkdown-html-number-figures:
outputFormat = opts_knit$get("rmarkdown.pandoc.to"); # determine the output format of the document
if( is.null(outputFormat) ) outputFormat = ""; # probably not run within knittr
capTabNo = 1; capFigNo = 1; # figure and table caption numbering, for HTML do it manually
#Function to add the Table Number
capTab = function(x)
{
  if(outputFormat == 'html'){
    x = paste0("**Table ",capTabNo,".** ",x,"")
    capTabNo <<- capTabNo + 1
  }; x
}

#Function to add the Figure Number
capFig = function(x, show_R_version=FALSE, show_package_versions=NULL, is_map=FALSE)
{
  if(outputFormat == 'html')
  {
    x <- paste0("**Figure ",capFigNo,".** ",x,"");
    if( show_R_version || (!is.null(show_package_versions) && length(show_package_versions) > 0) )
    {
      x <- paste0(x, " Figure generated using ");
      if( show_R_version ) x <- paste0(x, stringr::str_replace(R.version.string, stringr::fixed("R "), "[`R`](https://www.r-project.org/) "));
      if( !is.null(show_package_versions) && length(show_package_versions) > 0 )
      {
        x <- paste0(x, ifelse( show_R_version, " and ", " "));
        x <- paste0(x, ifelse( length(show_package_versions) > 1, "packages ", "package "));
        x <- paste0(x, paste0(vapply(show_package_versions, function(x) paste0("`",x,"`", " (version ", packageVersion(x),")"), character(1)), collapse=", "), ".");
      }
      if( is_map ) x <- paste0(x, " Maps are using public domain data from the [Natural Earth project](https://www.naturalearthdata.com/) as provided by the `R` package `maps`.");
    }
    capFigNo <<- capFigNo + 1;
  };
  x;
}

########## reproducibility ##########
set.seed(2137)
```

# Overview

**Research question.** Do speakers encode size iconically, and do listeners use those same
acoustic cues? Iconicity – a resemblance between linguistic form and meaning [@perniss_vigliocco_2014]
– has moved from a peripheral curiosity to a recognized organizing principle in language
[@cwiek_iconicity_2022; @cwiek_introduction_2025; @dingemanse_arbitrariness_2015], and size is one
of the most robust domains in which it surfaces: from the classic *mil/mal* contrast
[@sapir_study_1929] through the lexicon of English size adjectives [@winter_size_2021] and
cross-linguistic patterns [@blasi_soundmeaning_2016] to recent work on the acoustic and
articulatory bases of size–sound correspondences [@winter_size_2025].

**Expectations.** We expect both size and animacy to leave a trace on the two most-studied acoustic carriers of iconicity – vowel duration and f0 – and we expect this to hold bidirectionally: in production *and* perception. For size, the predicted direction is longer vowels and lower f0 for big referents, and listeners using those same cues to call "big". The f0 prediction follows the frequency code [@ohala_ethological_1984; @ohala_cross-language_1983], in which lower f0 indexes larger body size across vertebrates; the duration prediction follows recent treatments of iconicity as embedded in continuous acoustic and articulatory parameters rather than discrete phonemic features [@winter_size_2025; @fuchs_sounds_2022; @perlman_iconicity_2018]. The same auditory analogies are expected to guide listener categorization [@winter_size_2021; @perniss_vigliocco_2014; @schmidt_jabeen_dragon_2024]. For animacy we do not commit to a direction or a channel: animacy may enter as a modulator of the size effect (an interaction) or as an independent shift in the acoustic baseline (a main effect), or both, or neither. We test each of these openly. The literature on animacy and iconic production is mixed, so a single posterior that speaks to both modulation and main-effect possibilities is the safer reporting frame.

**Exploratory.** Other features – F1, F2, fricative duration, fricative SPL, fricative spectral peak, fricative centre of gravity, and the voice-quality measures CPP [@fraile_cepstral_2014; @hillenbrand_acoustic_1994] and H1–H2 [@chai_h1h2_2022] – are examined without a priori predictions. The pluripotential view of iconic features [cf. @akita_iconicity_2024; @winter_abralin_2021] motivates inspecting the broader signal rather than committing to a single channel; however, it also warrants the view that few features (f0 and duration) can correspond to many meanings.

**What we find, in one paragraph.** Speakers reliably lengthen vowels
and lower f0 for big referents (the production prediction, supported
on both channels). Listeners use f0 to judge size – lower f0 raises
the odds of a "big" judgment – but do not use vowel duration for the
same purpose (the perception prediction is supported for f0, not for
duration). The size effect does not differ between animate and
inanimate referents. Animate referents in this stimulus set were
produced with shorter vowels and higher f0 regardless of their size,
but this baseline shift does not survive the inclusion of a
separately-collected per-item cuteness covariate (§7.8); we therefore
report it descriptively without claiming animacy qua feature as the
carrier. Exploratory analyses recover a few additional channels worth
flagging – F1 lowers for big, CPP rises slightly for big, fricative
SPL is louder for big, and listeners use fricative duration and
fricative SPL as cues to size – without claiming these as robust
without replication.

This analysis uses the Bayesian framework with the `cmdstanr` backend. All German labels and column names from the original data files are translated to English in the cleaning step so the script and its outputs are readable internationally.

# Decisions log

- **Primary models** include the `size × animacy` interaction so a single posterior speaks to
  the size effect, the animacy main effect, and the modulation of size by animacy.
- **Random effects (production)**: `(1 + size_c | speaker) + (1 | item)`. 39 speakers, 36 items,
  ~1400 vowel rows / ~2470 fricative rows – comfortable.
- **Random effects (perception)**: `(1 | listener) + (1 | speaker) + (1 | item)`. 39 × 39 × 36,
  936 observations – varying slopes added per-model only if needed.
- **Coding**: sum contrasts `-0.5 / +0.5` for `size_c` and `animacy_c` so main effects remain
  interpretable in the presence of the interaction. Continuous predictors entering perception
  models are within-speaker z-scored – the listener-side coefficient then reads as "1 SD shift in
  *this speaker's* signal".
- **Priors**: weakly informative – `normal(0, 1)` on standardized fixed effects,
  `student_t(3, 0, 2.5)` on intercepts and SDs.
- **Outcome families**:
  - durations (vowel, fricative) → `lognormal()` (strictly positive, right-skewed)
  - f0 → `lognormal()` (positive, right-skewed; with sum-coded `size_c` the coefficient
    back-transforms cleanly to the ratio of f0 for big vs small)
  - formants (F1, F2) → `lognormal()` (positive, right-skewed)
  - CPP, H1–H2, fricative SPL, fricative peak frequency, fricative CoG → `gaussian()`
    (centred on their native scale; logging Hz outcomes overshoots the actual right-skew)
  - binary perception (`big_judgment`, `animate_judgment`) → `bernoulli(logit)`
  - 6-way `picked_category` → `categorical()` (kept available in §7.6 but not part
    of the reported analysis – see §7.6 for the reasoning)
- **Outliers**: not trimmed by default – the lognormal family absorbs the right tail. A
  diagnostic chunk flags extreme cases for inspection.
- **Caching**: fits saved to `models/`; `brms.file_refit = "on_change"` means re-knit only
  refits when the model spec changes.
- **Acoustic extraction**: vowel features via PraatSauce [@kirby_praatsauce_2025;
  @boersma_praat_2023]; fricative features via the Al-Tamimi spectral-analyses scripts
  [@al-tamimi_jalalal-tamimipraat-spectral-analyses-fricatives-stops_2022; @jongman_acoustic_2000].

# 1. Read raw data

`alle.csv` is comma-separated; the others are tab-separated. `out_praatSauce.tsv` has one
malformed row and uses `--undefined--` as a missing-value sentinel.

```{r read-raw}
alle_raw <- read_csv(file.path(data_dir, "alle.csv"),
                     show_col_types = FALSE) |>
  select(-1)        # drop the unnamed row-index column

praat_raw <- read_tsv(
  file.path(data_dir, "out_praatSauce.tsv"),
  na = c("", "NA", "--undefined--"),
  show_col_types = FALSE,
  col_types = cols(
    file        = col_character(),
    label       = col_character(),
    intervaldID = col_integer(),
    .default    = col_double()       # forces f0/F1/F2/CPP/H1H2/etc. to numeric
  )
) |>
  rename(intervalID = intervaldID) |>
  filter(!is.na(intervalID))         # drops the single malformed row (line 434)

fric_raw <- read_tsv(file.path(data_dir, "spectralResultsFricatives.tsv"),
                     show_col_types = FALSE)

vdur_raw <- read_tsv(file.path(data_dir, "vowelDuration.txt"),
                     show_col_types = FALSE) |>
  rename(file = FileName, vowel = digit, vowel_dur_ms = `Duration(ms)`)

cuteness_raw <- read_csv(file.path(data_dir, "data.csv"),
                         show_col_types = FALSE)
```

# 2. Translate German → English

Three layers: column names, categorical values, and (in a glossary) item words.
Stimulus item names are translated to `item_english`.

## 2.1 Column-name dictionary

```{r column-dict}
# Names of vector = original German column names; values = English replacements.
col_translations <- c(
  Gruppe              = "group",
  Bewertet            = "listener",
  Spricht             = "speaker",
  Belebtheit          = "animacy",
  Art                 = "category",
  Groesse             = "size",
  Exemplar            = "item",
  Kurz                = "item_code",
  Runde_Durchgang     = "round_trial",
  Geraten_Exemplar    = "picked_item",
  Geraten_Kurz        = "picked_item_code",
  Geraten_Belebtheit  = "picked_animacy",
  Geraten_Art         = "picked_category",
  Geraten_Groesse     = "picked_size",
  Exemplar_Korrekt    = "item_correct",
  Gruppe_Korrekt      = "group_correct",
  Belebtheit_Korrekt  = "animacy_correct",
  Art_Korrekt         = "category_correct",
  Groesse_Korrekt     = "size_correct"
)

translate_columns <- function(old_names) {
  new <- col_translations[old_names]
  ifelse(is.na(new), old_names, new)
}

alle <- alle_raw |> rename_with(translate_columns)
```

## 2.2 Value-label dictionaries

```{r value-translation}
alle <- alle |>
  mutate(
    across(c(size, picked_size),
           ~ case_when(.x == "Gross" ~ "big",
                       .x == "Klein" ~ "small",
                       TRUE ~ .x)),
    across(c(animacy, picked_animacy),
           ~ case_when(.x == "Belebt"   ~ "animate",
                       .x == "Unbelebt" ~ "inanimate",
                       TRUE ~ .x)),
    across(c(category, picked_category),
           ~ case_when(.x == "Vogel"      ~ "bird",
                       .x == "Saeugetier" ~ "mammal",
                       .x == "Reptil"     ~ "reptile",
                       .x == "Geraet"     ~ "appliance",
                       .x == "Fahrzeug"   ~ "vehicle",
                       .x == "Instrument" ~ "instrument",
                       TRUE ~ .x))
  )
```

## 2.3 Item-name glossary (German stimulus → English gloss)

```{r item-glossary}
item_glossary <- tribble(
  ~item_norm,            ~item_english,
  "buerodrucker",        "office.printer",
  "drucker",             "printer",
  "industriebackofen",   "industrial.oven",
  "backofen",            "oven",
  "armbanduhr",          "wristwatch",
  "wanduhr",             "wall.clock",
  "kirchenglocke",       "church.bell",
  "gloeckchen",          "little.bell",
  "kontrabass",          "double.bass",
  "violine",             "violin",
  "bodentrommel",        "floor.drum",
  "handtrommel",         "hand.drum",
  "grosserbagger",       "big.excavator",
  "kleinerbagger",       "small.excavator",
  "jumbojet",            "jumbo.jet",
  "kleinflugzeug",       "small.airplane",
  "suv",                 "SUV",
  "kleinwagen",          "small.car",
  "elefant",             "elephant",
  "kuh",                 "cow",
  "baer",                "bear",
  "kaninchen",           "rabbit",
  "aeffchen",            "little.monkey",
  "igel",                "hedgehog",
  "kondor",              "condor",
  "harpyie",             "harpy",
  "storch",              "stork",
  "blaumeise",           "blue.tit",
  "dompfaff",            "bullfinch",
  "spatz",               "sparrow",
  "anakonda",            "anaconda",
  "krokodil",            "crocodile",
  "komodowaran",         "komodo.dragon",
  "schildkroete",        "tortoise",
  "blindschleiche",      "slow.worm",
  "gecko",               "gecko"
)
```

# 3. Clean and harmonize

## 3.1 Parse production filenames

Format: `group + speaker_num _ trial_num _ entity _ sex + age` (e.g. `a1_01_buerodrucker_m31`).
Two filenames lack the `_sex+age` suffix (`d1_11_anakonda`, `i1_03_kontrabass`); we recover
speaker info from same-group siblings.

```{r parse-filenames}
parse_filename <- function(x) {
  x_clean <- x |> str_remove("\\.wav$") |> str_to_lower()
  m <- str_match(x_clean,
    "^([a-z])(\\d)_(\\d{2})_(.+?)_([mw])(\\d+)$")
  tibble(
    file_key      = x_clean,
    group         = m[, 2],
    speaker_num   = as.integer(m[, 3]),
    trial_num     = as.integer(m[, 4]),
    item_norm     = m[, 5],
    sex           = m[, 6],
    age           = as.integer(m[, 7])
  )
}

# Master file table: every unique production filename across the three production sources
master_files <- bind_rows(
  praat_raw |> distinct(file)     |> rename(file_orig = file),
  fric_raw  |> distinct(fileName) |> rename(file_orig = fileName),
  vdur_raw  |> distinct(file)     |> rename(file_orig = file)
) |>
  distinct(file_orig) |>
  mutate(parsed = map(file_orig, parse_filename)) |>
  unnest(parsed)

unparsed <- master_files |> filter(is.na(group))
cat("Files that did NOT parse:\n"); print(unparsed |> select(file_orig))
```

```{r recover-unparsed}
# Recover sex/age from same-group siblings (these participants appear elsewhere).
sibling_lookup <- master_files |>
  filter(!is.na(group)) |>
  distinct(group, speaker_num, sex, age)

manual_pat <- "^([a-z])(\\d)_(\\d{2})_(.+)$"
manual <- master_files |>
  filter(is.na(group)) |>
  mutate(file_key_lc = str_to_lower(file_orig) |> str_remove("\\.wav$")) |>
  transmute(
    file_orig,
    group       = str_match(file_key_lc, manual_pat)[, 2],
    speaker_num = as.integer(str_match(file_key_lc, manual_pat)[, 3]),
    trial_num   = as.integer(str_match(file_key_lc, manual_pat)[, 4]),
    item_norm   = str_match(file_key_lc, manual_pat)[, 5],
    file_key    = file_key_lc
  ) |>
  left_join(sibling_lookup, by = c("group", "speaker_num"))

master_files <- master_files |>
  filter(!is.na(group)) |>
  bind_rows(manual) |>
  distinct(file_key, .keep_all = TRUE)   # dedupe .wav / no-extension variants

stopifnot(sum(is.na(master_files$group)) == 0)
glimpse(master_files)
```

## 3.2 Normalize item names (fix typos, harmonize spelling)

Production filenames and the perception `item` column use different conventions, plus a handful
of typos (`harypie`/`harpyie`, `industrienbackofen`/`industriebackofen`,
`schildkorete`/`schildkroete`). We normalize both sides to the same lowercase, no-spaces form.

```{r normalize-items}
normalize_item <- function(x) {
  x |>
    str_to_lower() |>
    str_trim() |>
    str_remove_all("\\s+") |>
    str_replace("^harypie$",            "harpyie") |>
    str_replace("^industrienbackofen$", "industriebackofen") |>
    str_replace("^schildkorete$",       "schildkroete")
}

master_files <- master_files |>
  mutate(item_norm = normalize_item(item_norm))

alle <- alle |>
  mutate(item_norm = normalize_item(item))

# Verify item-name correspondence after normalization
prod_items <- master_files |> distinct(item_norm) |> pull(item_norm) |> sort()
perc_items <- alle |> distinct(item_norm) |> pull(item_norm) |> sort()

cat("Items in production but not perception:\n");
print(setdiff(prod_items, perc_items))
cat("\nItems in perception but not production:\n");
print(setdiff(perc_items, prod_items))
```

## 3.3 Parse perception speaker / listener codes

```{r parse-perception-codes}
parse_speaker_code <- function(s) {
  m <- str_match(s, "^([a-z])(\\d)_([mw])_(\\d+)$")
  tibble(
    .grp     = m[, 2],
    .spknum  = as.integer(m[, 3]),
    .sex     = m[, 4],
    .age     = as.integer(m[, 5])
  )
}

alle <- alle |>
  bind_cols(parse_speaker_code(alle$speaker)  |> rename_with(~ paste0("spk", .))) |>
  bind_cols(parse_speaker_code(alle$listener) |> rename_with(~ paste0("lst", .))) |>
  separate(round_trial, into = c("round", "trial_num"),
           sep = "_", remove = FALSE, convert = TRUE) |>
  rename(speaker_num  = spk.spknum,
         speaker_sex  = spk.sex,
         speaker_age  = spk.age,
         listener_sex = lst.sex,
         listener_age = lst.age) |>
  select(-spk.grp, -lst.grp)

# Sanity: round must equal speaker_num
stopifnot(all(alle$round == alle$speaker_num))

# Binary outcomes for downstream models
alle <- alle |>
  mutate(
    big_judgment      = as.integer(picked_size      == "big"),
    animate_judgment  = as.integer(picked_animacy   == "animate"),
    portrayed_big     = as.integer(size             == "big"),
    portrayed_animate = as.integer(animacy          == "animate")
  )
```

```{r mismatch-diagnostic}
# Quality-control trace: trials where the production filename and the perception
# sheet disagree about which item was used. Per speaker verification, the
# perception sheet is authoritative – these are listed for documentation only.

trial_item_check <- master_files |>
  distinct(group, speaker_num, trial_num, item_norm) |>
  rename(item_from_filename = item_norm) |>
  inner_join(
    alle |> distinct(group, speaker_num, trial_num, item_norm) |>
      rename(item_from_perception = item_norm),
    by = c("group", "speaker_num", "trial_num")
  ) |>
  filter(item_from_filename != item_from_perception)

if (nrow(trial_item_check) > 0) {
  cat("Trial-level item-name disagreements between sources",
      "(perception sheet is authoritative; production filename ignored for item identity):\n")
  print(trial_item_check)
}
```

## 3.4 Cuteness ratings (separate rater sample)

These ratings come from a different sample of participants than the
production/perception experiment – they only rated stimulus pictures for
cuteness on (we assume) a 1–7 scale. We exclude participants from abroad, then summarize to one mean cuteness score per stimulus item.
The per-item score is the only level used downstream, because the cuteness
raters cannot be matched 1:1 to listeners or speakers.

```{r cuteness-clean}
cuteness_clean <- cuteness_raw |>
  select(-gender_other, -COUNTRY, -CLIENT_start, -TIME_start, -TIME_end, -TIME_total) |>
  filter(if_all(everything(), ~ !is.na(.))) |>
  filter(!country %in% c("albania", "Albania", "France, India, England", "France")) |>
  mutate(gender = case_when(
    gender == 1 ~ "female", 
    gender == 2 ~ "male", 
    TRUE ~ NA_character_))

cat("Cuteness raters retained:", nrow(cuteness_clean),
    "of", nrow(cuteness_raw), "\n")
```

```{r cuteness-map}
# Map cuteness column names to item_english (§2.3 glossary).
# Six columns (heron, horse, frog, washingmachine, toaster, boat) are
# fillers in the cuteness study with no production counterpart -- dropped.
cuteness_map <- tribble(
  ~col,                 ~item_english,
  "condor",             "condor",
  "harpyie",            "harpy",
  "stork",              "stork",
  "finch",              "bullfinch",       
  "tit",                "blue.tit",
  "sparrow",            "sparrow",
  "cow",                "cow",
  "elephant",           "elephant",
  "bear",               "bear",
  "monkey",             "little.monkey",
  "hedgehog",           "hedgehog",
  "bunny",              "rabbit",
  "crocodile",          "crocodile",
  "comododragon",       "komodo.dragon",
  "anaconda",           "anaconda",
  "gecko",              "gecko",
  "grasssnake",         "slow.worm",       
  "tortoise",           "tortoise",
  "doublebass",         "double.bass",
  "bigdrum",            "floor.drum",
  "bigbell",            "church.bell",
  "violin",             "violin",
  "smalldrum",          "hand.drum",
  "smallbell",          "little.bell",
  "wallclock",          "wall.clock",
  "industrialoven",     "industrial.oven",
  "industrialprinter",  "office.printer",  
  "watch",              "wristwatch",
  "oven",               "oven",
  "printer",            "printer",
  "SUV",                "SUV",
  "jumbojet",           "jumbo.jet",
  "bigexcavator",       "big.excavator",
  "car",                "small.car",
  "plane",              "small.airplane",
  "smallexcavator",     "small.excavator"
)

# Per-rater long form
cuteness_long <- cuteness_clean |>
  select(participant, all_of(cuteness_map$col)) |>
  pivot_longer(-participant, names_to = "col", values_to = "cuteness") |>
  inner_join(cuteness_map, by = "col") |>
  select(-col)

# Per-item summary: mean, SD, n_raters.
# cuteness_z is the z-scored mean (across the 36 items) used by the models.
item_cuteness <- cuteness_long |>
  group_by(item_english) |>
  summarise(cuteness_mean = mean(cuteness, na.rm = TRUE),
            cuteness_sd   = sd(cuteness,   na.rm = TRUE),
            n_raters      = sum(!is.na(cuteness)),
            .groups = "drop") |>
  mutate(cuteness_z = as.numeric(scale(cuteness_mean)))
```

```{r cuteness-coverage}
# Coverage diagnostic: are any of the 36 stimulus items missing a rating?
items_in_study <- alle |> distinct(item_norm) |>
  left_join(item_glossary, by = "item_norm") |>
  pull(item_english)

missing_cute <- setdiff(items_in_study, item_cuteness$item_english)
cat("Stimulus items with no cuteness score:",
    if (length(missing_cute) == 0) "none" else paste(missing_cute, collapse = ", "),
    "\n")

cat("Rater sample size, per item:\n")
print(item_cuteness |> count(n_raters))
```

```{r cuteness-by-item-prep}
# Per-item table enriched with size, animacy, AND the 6-way category.
# The category column lives in `alle`, so we join through item_glossary.
item_cute_full <- item_cuteness |>
  left_join(
    alle |> distinct(item_norm, size, animacy, category) |>
      left_join(item_glossary, by = "item_norm") |>
      select(item_english, size, animacy, category),
    by = "item_english"
  ) |>
  mutate(
    cuteness_se = cuteness_sd / sqrt(n_raters),
    # Order categories: animals first, objects second; within each block,
    # ordered roughly by intuition (mammals > birds > reptiles; appliances,
    # instruments, vehicles)
    category = factor(category,
                      levels = c("mammal", "bird", "reptile",
                                 "appliance", "instrument", "vehicle"))
  )
```

```{r cuteness-descriptive, fig.height=5.5, fig.width=9, fig.cap=capFig("Per-item cuteness by size and animacy. Each labelled point is one of the 36 stimulus items; black bars are cell means. The animacy gap is the largest visual feature (~1.3 points). Within animals, small items sit visibly above big items (rabbit and hedgehog top the chart; cow and anaconda anchor the bottom); within objects the size effect is much smaller and noisier. Note elephant as a high outlier among big animals, and small.car as the floor among objects.")}
# Build the enriched per-item table (item_cuteness + size/animacy/category).
# Created here so this chunk is self-contained; the later viz chunks
# (cuteness-ranked, cuteness-by-category) reuse the object.
item_cute_full <- item_cuteness |>
  left_join(
    alle |> distinct(item_norm, size, animacy, category) |>
      left_join(item_glossary, by = "item_norm") |>
      select(item_english, size, animacy, category),
    by = "item_english"
  ) |>
  mutate(
    cuteness_se = cuteness_sd / sqrt(n_raters),
    category    = factor(category,
                         levels = c("mammal", "bird", "reptile",
                                    "appliance", "instrument", "vehicle"))
  )

# Pre-jitter so geom_text_repel and geom_point share the same x positions.
set.seed(2137)
plot_dat <- item_cute_full |>
  mutate(
    size_num = as.numeric(factor(size, levels = c("big", "small"))),
    x        = size_num + runif(n(), -0.09, 0.09)
  )

# Cell means as a tiny tibble we draw as crossbars
cell_means <- item_cute_full |>
  group_by(animacy, size) |>
  summarise(cm = mean(cuteness_mean), .groups = "drop") |>
  mutate(size_num = as.numeric(factor(size, levels = c("big", "small"))))

ggplot(plot_dat, aes(x = x, y = cuteness_mean)) +
  geom_segment(data = cell_means,
               aes(x = size_num - 0.28, xend = size_num + 0.28,
                   y = cm, yend = cm),
               colour = "black", linewidth = 0.7,
               inherit.aes = FALSE) +
  geom_text(data = cell_means,
            aes(x = size_num, y = 7.05, label = sprintf("M = %.2f", cm)),
            colour = "grey25", size = 3.1, fontface = "italic",
            inherit.aes = FALSE) +
  geom_point(aes(colour = animacy), size = 2.6, alpha = 0.9) +
  ggrepel::geom_text_repel(
    aes(label = item_english, colour = animacy),
    size = 3.0, max.overlaps = Inf,
    box.padding = 0.32, point.padding = 0.18,
    min.segment.length = 0,
    segment.alpha = 0.45, segment.size = 0.25,
    force = 1.4, force_pull = 0.6,
    seed = 2137,
    show.legend = FALSE
  ) +
  facet_wrap(~ animacy) +
  scale_x_continuous(
    breaks = c(1, 2), labels = c("big", "small"),
    limits = c(0.4, 2.6)
  ) +
  scale_y_continuous(limits = c(1, 7.25), breaks = 1:7) +
  scale_colour_manual(values = colorBlindBlack8[c(3, 4)]) +
  labs(
    x = NULL, y = "Mean cuteness rating (1–7)",
    title    = "Per-item cuteness by size and animacy"
  ) +
  theme(
    legend.position  = "none",
    panel.spacing.x  = unit(1.2, "lines"),
    strip.text       = element_text(face = "bold", size = 11)
  )
```


```{r cuteness-ranked, fig.height=8, fig.cap=capFig("Per-item mean cuteness rating across the 36 stimulus items, ranked. Points are item means; horizontal bars are ±1 SE across the 16 raters. Animals (animate) dominate the top of the ranking and objects (inanimate) the bottom – the structural confound between cuteness and animacy is visible at a glance.")}
ggplot(item_cute_full,
       aes(x = cuteness_mean, y = fct_reorder(item_english, cuteness_mean),
           colour = animacy, shape = size)) +
  geom_segment(aes(xend = 1, yend = fct_reorder(item_english, cuteness_mean)),
               colour = "grey85", linewidth = 0.3) +
  geom_pointrange(aes(xmin = cuteness_mean - cuteness_se,
                      xmax = cuteness_mean + cuteness_se),
                  size = 0.4, fatten = 3) +
  scale_colour_manual(values = colorBlindBlack8[c(3, 4)]) +
  scale_shape_manual(values = c(small = 16, big = 17)) +
  scale_x_continuous(limits = c(1, 7), breaks = 1:7) +
  labs(x = "Mean cuteness rating (1–7), ± 1 SE across raters",
       y = NULL,
       colour = "Animacy", shape = "Size",
       title = "Per-item cuteness, ranked")
```

```{r cuteness-by-category, fig.height=4, fig.cap=capFig("Per-item cuteness by stimulus category. Each dot is one item; black bar is the category mean. Within-animacy structure is visible: mammals are rated cutest, reptiles least cute among animals; among objects, instruments rate higher than appliances or vehicles.")}
ggplot(item_cute_full, aes(x = category, y = cuteness_mean, colour = animacy)) +
  geom_jitter(width = 0.18, height = 0, size = 2.6, alpha = 0.85) +
  stat_summary(fun = mean, geom = "crossbar",
               width = 0.55, linewidth = 0.3, colour = "black",
               fatten = 0) +
  scale_colour_manual(values = colorBlindBlack8[c(3, 4)]) +
  scale_y_continuous(limits = c(1, 7), breaks = 1:7) +
  labs(x = NULL, y = "Mean cuteness rating",
       colour = "Animacy",
       title = "Cuteness by stimulus category")
```


## 3.5 Cuteness check

```{r m-cuteness-prep}
# Attach size/animacy to the rater-level ratings and sum-code (+0.5 = big, animate),
# matching the coding used everywhere else in the script.
cuteness_model_dat <- cuteness_long |>
  left_join(item_cute_full |> select(item_english, size, animacy),
            by = "item_english") |>
  filter(!is.na(size), !is.na(animacy)) |>
  mutate(
    size_c    = if_else(size    == "big",     0.5, -0.5),
    animacy_c = if_else(animacy == "animate", 0.5, -0.5)
  )
```

```{r m-cuteness-fit, results='hide'}
# Multilevel Gaussian model on the 1–7 ratings: effects come out in rating points,
# which match the descriptive "~1.3-point animacy gap" language. By-rater and
# by-item random intercepts handle the clustering the pooled Wilcoxon ignored.
# Priors are weakly informative, scaled to the 1–7 outcome.
m_cuteness <- brm(
  bf(cuteness ~ size_c * animacy_c + (1 | participant) + (1 | item_english)),
  data    = cuteness_model_dat,
  family  = gaussian(),
  prior   = c(prior(normal(0, 2.5),        class = "b"),
              prior(student_t(3, 4, 2.5),  class = "Intercept"),
              prior(student_t(3, 0, 2.5),  class = "sd"),
              prior(student_t(3, 0, 2.5),  class = "sigma")),
  chains  = 4, iter = 8000, warmup = 4000, cores = 4,
  control = list(adapt_delta = 0.95, max_treedepth = 12),
  seed    = 2137,
  file    = paste0(models, "m_cuteness")
)
```

```{r m-cuteness-report}
summary(m_cuteness)

# Same reporting idiom as the rest of the paper: median, 95% HDI, prob. of direction.
describe_posterior(m_cuteness, ci = 0.95, ci_method = "hdi",
                   test = "pd", centrality = "median", effects = "fixed")

# Optional PPC, consistent with your other models:
pp_check(m_cuteness, ndraws = 100) + ggtitle("PPC – cuteness ~ size * animacy")
```

The images were generated in a uniform "cute" style so that cuteness would not vary with the experimental factors, but a style prompt is no guarantee, so we check directly whether the ratings still line up with size and animacy. We model the raw rater-level ratings – every rater × item, 576 in all – rather than the per-item means, with random intercepts for both rater and item: raters differ in how generously they use the scale, and items differ from one another beyond their size and animacy, and a multilevel model absorbs both kinds of variation instead of letting them inflate the precision of the size and animacy estimates (it is also why a single pooled test would overstate the evidence here). The outcome is treated as continuous, so the coefficients are in rating points on the 1–7 scale; size and animacy are sum-coded (+0.5 = big, animate), so each coefficient is the difference between its two levels.

The check confirms the confound the covariate is meant to handle. Animate referents are rated about 1.5 points cuter than inanimate ones (β = 1.53, 95% CrI [1.26, 1.80]), and small referents about half a point cuter than big ones (β = −0.47, 95% CrI [−0.74, −0.20]); both intervals exclude zero. The size × animacy interaction leans the same way as the descriptive plots – the size effect on cuteness is a little larger for animals than for objects – but its interval crosses zero (β = −0.48, 95% CrI [−1.01, 0.07]), so we read it as a hint rather than a result. All chains converged (R-hat = 1.00). Because cuteness covaries with both factors, it enters the production and perception models as a per-item covariate; whether it actually accounts for the animacy main effect is taken up in §7.8.

# 4. Build the long-format production tables

Production data are in **long format**: one row per acoustic measurement (vowel or fricative
token). Per-trial summaries used in perception models are derived from these.

Files with <3 vowels reflect real audio/production issues (per speaker verification); rows with NA vowel_dur_ms are dropped per-vowel during modelling – no global filtering needed.

```{r build-prod-long}
# ---- vowel duration: one row per vowel
vdur_long <- vdur_raw |>
  mutate(file_key = str_to_lower(file)) |>
  left_join(master_files |> select(file_key, group, speaker_num, trial_num,
                                   item_norm, sex, age),
            by = "file_key") |>
  filter(!is.na(group))

# ---- praatSauce: one row per vowel, with f0 / F1 / F2 / CPP / H1-H2.
# H1-H2 = A1c - A2c per the readme.
praat_long <- praat_raw |>
  mutate(file_key = str_to_lower(file) |> str_remove("\\.wav$"),
         H1H2     = H1c - H2c) |>
  rename(vowel = label) |>
  left_join(master_files |> select(file_key, group, speaker_num, trial_num,
                                   item_norm, sex, age),
            by = "file_key") |>
  filter(!is.na(group))

# Join vowel duration onto praat (per file × vowel)
vowel_long <- praat_long |>
  left_join(vdur_long |> select(file_key, vowel, vowel_dur_ms),
            by = c("file_key", "vowel"))

# ---- Fricatives: one row per fricative token
fric_long <- fric_raw |>
  mutate(file_key = str_to_lower(fileName) |> str_remove("\\.wav$")) |>
  rename(fric_dur_ms  = durationMS,
         fric_SPL_dB  = levelMWholedB,    # whole-segment, mid-band intensity
         fric_peak_Hz = peakFreqHzMid,
         fric_CoG_Hz  = cogMid) |>
  filter(phoneme %in% c("C", "f", "z", "S")) |>   # drop one anomalous "´z" record
  left_join(master_files |> select(file_key, group, speaker_num, trial_num,
                                   item_norm, sex, age),
            by = "file_key") |>
  filter(!is.na(group)) |>
  select(file_key, phoneme, fric_dur_ms, fric_SPL_dB, fric_peak_Hz, fric_CoG_Hz,
         group, speaker_num, trial_num, item_norm, sex, age)
```

## 4.1 Attach experimental factors and item glosses to production rows

```{r attach-factors}
trial_factors <- alle |>
  distinct(group, speaker_num, trial_num, item_norm,
           animacy, category, size, item_code, item) |>
  rename(item_label = item)

speaker_id <- master_files |>
  distinct(group, speaker_num, sex, age) |>
  mutate(speaker = paste0(group, speaker_num, "_", sex, "_", age))

vowel_long <- vowel_long |>
  select(-item_norm) |>                       # filename's label is unreliable; perception wins
  left_join(trial_factors, by = c("group", "speaker_num", "trial_num")) |>
  left_join(speaker_id,    by = c("group", "speaker_num", "sex", "age")) |>
  left_join(item_glossary, by = "item_norm")

fric_long <- fric_long |>
  select(-item_norm) |>
  left_join(trial_factors, by = c("group", "speaker_num", "trial_num")) |>
  left_join(speaker_id,    by = c("group", "speaker_num", "sex", "age")) |>
  left_join(item_glossary, by = "item_norm")

# Per-item cuteness covariate (mean across the 17 retained raters; z-scored over items)
vowel_long <- vowel_long |>
  left_join(item_cuteness |> select(item_english, cuteness_mean, cuteness_z),
            by = "item_english")

fric_long <- fric_long |>
  left_join(item_cuteness |> select(item_english, cuteness_mean, cuteness_z),
            by = "item_english")

cat("vowel_long rows missing factors:", sum(is.na(vowel_long$size)), "\n")
cat("fric_long  rows missing factors:", sum(is.na(fric_long$size)),  "\n")
```

## 4.2 Build the perception dataframe with speaker-level acoustic predictors

For each (group, speaker, trial), summarize the speaker's acoustic output for that production –
mean across vowels for vocalic features, mean across fricatives for fricative features. Predictors
are then within-speaker z-scored so the perception coefficient reads as "what does +1 SD of *this
speaker's* signal do to listener choice?".

```{r build-perception}
# Per-trial vowel summaries
trial_vowel <- vowel_long |>
  group_by(group, speaker_num, trial_num) |>
  summarise(
    vowel_dur_ms_mean = mean(vowel_dur_ms, na.rm = TRUE),
    f0_mean           = mean(f0,           na.rm = TRUE),
    F1_mean           = mean(F1,           na.rm = TRUE),
    F2_mean           = mean(F2,           na.rm = TRUE),
    CPP_mean          = mean(CPP,          na.rm = TRUE),
    H1H2_mean         = mean(H1H2,         na.rm = TRUE),
    .groups = "drop"
  )

# Per-trial fricative summaries
trial_fric <- fric_long |>
  group_by(group, speaker_num, trial_num) |>
  summarise(
    fric_dur_ms_mean  = mean(fric_dur_ms,  na.rm = TRUE),
    fric_SPL_dB_mean  = mean(fric_SPL_dB,  na.rm = TRUE),
    fric_peak_Hz_mean = mean(fric_peak_Hz, na.rm = TRUE),
    fric_CoG_Hz_mean  = mean(fric_CoG_Hz,  na.rm = TRUE),
    .groups = "drop"
  )

zscale <- function(x) as.numeric(scale(x))

perc <- alle |>
  left_join(trial_vowel,   by = c("group", "speaker_num", "trial_num")) |>
  left_join(trial_fric,    by = c("group", "speaker_num", "trial_num")) |>
  left_join(item_glossary, by = "item_norm") |>
  group_by(speaker) |>
  mutate(across(c(vowel_dur_ms_mean, f0_mean, F1_mean, F2_mean, CPP_mean, H1H2_mean,
                  fric_dur_ms_mean, fric_SPL_dB_mean, fric_peak_Hz_mean, fric_CoG_Hz_mean),
                zscale, .names = "{.col}_z")) |>
  ungroup()

perc <- perc |>
  left_join(item_cuteness |> select(item_english, cuteness_mean, cuteness_z),
            by = "item_english")

cat("Perception rows without production audio:",
    sum(is.na(perc$vowel_dur_ms_mean)), "of", nrow(perc), "\n")
```

## 4.3 Sum-coding for size and animacy

```{r factor-coding}
# Helper: 2-level sum-coded contrast (-0.5 / +0.5)
sum_code_2 <- function(x, levels) {
  x <- factor(x, levels = levels)
  as.numeric(x) - 1.5      # level 1 -> -0.5, level 2 -> +0.5
}

# Production-side
vowel_long <- vowel_long |>
  mutate(
    size_c    = sum_code_2(size,    c("small",     "big")),       # +0.5 = big
    animacy_c = sum_code_2(animacy, c("inanimate", "animate"))    # +0.5 = animate
  )

fric_long <- fric_long |>
  mutate(
    size_c    = sum_code_2(size,    c("small",     "big")),
    animacy_c = sum_code_2(animacy, c("inanimate", "animate"))
  )

# Perception-side
perc <- perc |>
  mutate(
    size_c    = sum_code_2(size,    c("small",     "big")),
    animacy_c = sum_code_2(animacy, c("inanimate", "animate"))
  )
```

# 5. Exploratory data analysis

## 5.1 Sample sizes

```{r eda-counts}
cat("Vowel rows:        ", nrow(vowel_long), "\n")
cat("Fricative rows:    ", nrow(fric_long),  "\n")
cat("Perception rows:   ", nrow(perc),       "\n")
cat("Speakers:          ", n_distinct(perc$speaker),   "\n")
cat("Listeners:         ", n_distinct(perc$listener),  "\n")
cat("Items:             ", n_distinct(perc$item_english), "\n")

perc |> tabyl(animacy, size, category) |> print()
```

## 5.2 Distributions of acoustic features

```{r eda-distributions, fig.height=6, fig.cap=capFig("Vocalic features by size and animacy. Median lines inside violins.")}
acoustic_long <- vowel_long |>
  select(speaker, item_english, vowel, size, animacy,
         vowel_dur_ms, f0, F1, F2, CPP, H1H2) |>
  pivot_longer(c(vowel_dur_ms, f0, F1, F2, CPP, H1H2),
               names_to = "feature", values_to = "value")

ggplot(acoustic_long, aes(x = size, y = value, fill = animacy)) +
  geom_violin(draw_quantiles = 0.5, alpha = 0.7) +
  facet_wrap(~ feature, scales = "free_y") +
  scale_fill_manual(values = colorBlindBlack8[c(3, 4)]) +
  labs(title = "Vocalic features by size and animacy",
       x = NULL, y = NULL)
```

```{r eda-fricatives, fig.height=4, fig.cap=capFig("Fricative features by size and animacy.")}
fric_long |>
  pivot_longer(c(fric_dur_ms, fric_SPL_dB, fric_peak_Hz, fric_CoG_Hz),
               names_to = "feature", values_to = "value") |>
  ggplot(aes(x = size, y = value, fill = animacy)) +
  geom_violin(draw_quantiles = 0.5, alpha = 0.7) +
  facet_wrap(~ feature, scales = "free_y") +
  scale_fill_manual(values = colorBlindBlack8[c(3, 4)]) +
  labs(title = "Fricative features by size and animacy",
       x = NULL, y = NULL)
```

## 5.3 Outlier flags (vowel duration)

```{r outlier-flags}
vowel_long |>
  group_by(vowel) |>
  mutate(z = scale(log(vowel_dur_ms))[, 1]) |>
  ungroup() |>
  filter(abs(z) > 3) |>
  select(file_key, vowel, vowel_dur_ms, z, size, animacy) |>
  arrange(desc(abs(z)))
```

Extreme productions are retained: speaker verification confirmed these are genuine iconic productions (not measurement errors). Trimming would discard signal. The lognormal family on duration accommodates the right tail; PPCs are the diagnostic that matters.

## 5.4 Perception accuracy snapshot

```{r eda-accuracy}
perc |>
  summarise(
    size_acc      = mean(size_correct),
    animacy_acc   = mean(animacy_correct),
    category_acc  = mean(category_correct),
    item_acc      = mean(item_correct)
  )
```

```{r eda-perc-by-cell, fig.height=4, fig.cap=capFig("Proportion of 'big' judgments by portrayed size and animacy of the referent.")}
perc |>
  group_by(animacy, size) |>
  summarise(prop_big_judgment = mean(big_judgment), .groups = "drop") |>
  ggplot(aes(x = size, y = prop_big_judgment, fill = animacy)) +
  geom_col(position = position_dodge()) +
  geom_hline(yintercept = 0.5, linetype = 2) +
  scale_fill_manual(values = colorBlindBlack8[c(3, 4)]) +
  labs(y = "P(\"big\" judgment)",
       title = "Listener \"big\" judgment by portrayed size and animacy")
```

# 6. Primary analyses – duration and f0

The primary models test our expectations directly: size and animacy should each leave a trace
on vowel duration and f0, in both production and perception. Each fit includes `size_c`,
`animacy_c`, and their interaction so that a single posterior can answer three questions at
once: (i) is there a size effect in the expected direction, (ii) does animacy shift the
outcome independent of size, and (iii) does animacy modulate the size effect? Fitting all three
together avoids the multiple-testing pitfalls of separate models. For size the predicted
direction is fixed in advance (longer for big, lower f0 for big); for animacy we keep
direction and channel (interaction vs. main effect) open.

## 6.1 A reusable fitter

```{r helper-fit}
default_priors <- function(family = "gaussian") {
  switch(family,
    gaussian  = c(prior(normal(0, 100),        class = "b"),
                  prior(student_t(3, 0, 100),  class = "Intercept"),
                  prior(student_t(3, 0, 100),  class = "sd"),
                  prior(student_t(3, 0, 100),  class = "sigma")),
    lognormal = c(prior(normal(0, 1.5),        class = "b"),
                  prior(student_t(3, 5, 5),    class = "Intercept"),
                  prior(student_t(3, 0, 2),    class = "sd"),
                  prior(student_t(3, 0, 2),    class = "sigma")),
    bernoulli = c(prior(normal(0, 5),          class = "b"),
                  prior(student_t(3, 0, 10),   class = "Intercept"),
                  prior(student_t(3, 0, 5),    class = "sd"))
  )
}

fit_brm <- function(formula, data, family = gaussian(), file = NULL,
                    priors = NULL, ...) {
  fam_name <- family$family
  if (is.null(priors)) priors <- default_priors(fam_name)
  # Inform the user whether we're hitting the cache or rebuilding
  if (!is.null(file)) {
    rds <- if (endsWith(file, ".rds")) file else paste0(file, ".rds")
    if (file.exists(rds)) {
      message("[fit_brm] Cache exists at '", basename(rds),
              "' – brms will reuse it if formula/data/prior match.")
    } else {
      message("[fit_brm] No cache at '", basename(rds), "' – fitting from scratch.")
    }
  }
  brm(formula = formula, data = data, family = family,
      prior = priors,
      chains = 4, iter = 8000, warmup = 4000, cores = 4,
      control = list(adapt_delta = 0.95, max_treedepth = 12),
      seed   = 2137,
      file   = file,
      refresh = 100,
      ...)
}
```

## 6.2 Production: vowel duration

```{r m-vowel-dur, results='hide'}
m_vdur <- fit_brm(
  bf(vowel_dur_ms ~ size_c * animacy_c + vowel +
       (1 + size_c | speaker) + (1 | item_english)),
  data   = vowel_long |> filter(!is.na(vowel_dur_ms)),
  family = lognormal(),
  file   = paste0(models, "m_vdur")
)
```

```{r m-vowel-dur-ppc, fig.cap=capFig("Posterior predictive check for vowel duration (lognormal).")}
summary(m_vdur)
pp_check(m_vdur, ndraws = 100) + ggtitle("PPC – vowel duration")
```

## 6.3 Production: f0

f0 is modelled with `lognormal()` because it is strictly positive and right-skewed, especially
once male and female speakers are pooled. The covariate `sex` is included because male and
female speakers differ by roughly 80–100 Hz on average; without `sex` the random intercept
absorbs that difference but the fixed-effect intercept becomes harder to interpret. The
lognormal choice has a pleasant analytical side-effect: with the sum-coded `size_c` predictor
(`-0.5` for small, `+0.5` for big), the coefficient is exactly `log(f0_big / f0_small)`, so
`exp(β)` is the multiplicative ratio of speaker f0 for big versus small productions – a clean
quantity for the frequency-code prediction, which itself is a *ratio* claim (lower f0 indexes
larger body size across vertebrates) rather than a Hz claim. The prediction – lower f0 for
big referents – is the central frequency-code expectation [@ohala_ethological_1984;
@ohala_cross-language_1983].

```{r m-f0, results='hide'}
m_f0 <- fit_brm(
  bf(f0 ~ size_c * animacy_c + sex + vowel +
       (1 + size_c | speaker) + (1 | item_english)),
  data   = vowel_long |> filter(!is.na(f0)),
  family = lognormal(),
  file   = paste0(models, "m_f0")
)
```

```{r m-f0-ppc, fig.cap=capFig("Posterior predictive check for f0 (lognormal, with sex covariate).")}
summary(m_f0)
pp_check(m_f0, ndraws = 100) + ggtitle("PPC – f0")
```

## 6.4 Perception: vowel duration → "big" judgment

```{r m-perc-dur, results='hide'}
m_perc_dur <- fit_brm(
  bf(big_judgment ~ vowel_dur_ms_mean_z * animacy_c +
       (1 | listener) + (1 | speaker) + (1 | item_english)),
  data   = perc |> filter(!is.na(vowel_dur_ms_mean_z)),
  family = bernoulli(),
  file   = paste0(models, "m_perc_dur")
)
```

```{r m-perc-dur-ppc, fig.cap=capFig("Posterior predictive check for perception of duration (Bernoulli).")}
summary(m_perc_dur)
pp_check(m_perc_dur, ndraws = 100) + ggtitle("PPC – duration perception")
```

## 6.5 Perception: f0 → "big" judgment

```{r m-perc-f0, results='hide'}
m_perc_f0 <- fit_brm(
  bf(big_judgment ~ f0_mean_z * animacy_c +
       (1 | listener) + (1 | speaker) + (1 | item_english)),
  data   = perc |> filter(!is.na(f0_mean_z)),
  family = bernoulli(),
  file   = paste0(models, "m_perc_f0")
)
```

```{r m-perc-f0-ppc, fig.cap=capFig("Posterior predictive check for perception of f0 (Bernoulli).")}
summary(m_perc_f0)
pp_check(m_perc_f0, ndraws = 100) + ggtitle("PPC – f0 perception")
```

## 6.6 Reporting helpers

Two small helpers do the heavy lifting for everything that follows. `report_model()` prints a
posterior table (medians and 95% HDIs), a posterior predictive check, and a half-eye plot of
the fixed-effect posteriors for a given model. `report_effect()` zeroes in on a single
coefficient – useful when we want the back-transformed quantity (percent change for lognormal,
odds ratio for Bernoulli) and the one-sided posterior probability for a directional claim.

```{r report-helper}
report_model <- function(model, label = "model", terms_of_interest = NULL) {
  cat("\n========== ", label, " ==========\n")

  ds <- describe_posterior(model, ci = 0.95, ci_method = "hdi",
                           test = c("pd"), centrality = "median",
                           effects = "fixed")
  print(ds)

  print(pp_check(model, ndraws = 100) + ggtitle(paste("PPC –", label)))

  draws <- as_draws_df(model) |>
    select(starts_with("b_"), -matches("^b_Intercept$")) |>
    pivot_longer(everything(), names_to = "term", values_to = "value") |>
    mutate(term = str_remove(term, "^b_"))

  if (!is.null(terms_of_interest))
    draws <- draws |> filter(term %in% terms_of_interest)

  p <- ggplot(draws, aes(x = value, y = term)) +
    stat_halfeye(.width = c(0.66, 0.95)) +
    geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
    labs(title = paste("Posterior –", label), x = "estimate", y = NULL)
  print(p)

  invisible(ds)
}

report_effect <- function(model, coef_name,
                          direction = c("auto", "positive", "negative"),
                          label = NULL) {
  direction   <- match.arg(direction)
  family_name <- family(model)$family
  if (is.null(label)) label <- paste0(deparse(substitute(model)), " :: ", coef_name)

  # ---- Posterior draws of the coefficient on the model's link scale
  beta <- as_draws_df(model)[[paste0("b_", coef_name)]]
  if (is.null(beta)) stop("Coefficient '", coef_name, "' not in model.")

  link_hdi <- median_hdi(beta, .width = 0.95)

  # ---- Decide direction for the one-sided hypothesis test
  if (direction == "auto") {
    direction <- if (median(beta) > 0) "positive" else "negative"
  }
  hyp_str <- paste(coef_name, if (direction == "positive") "> 0" else "< 0")
  hyp     <- hypothesis(model, hyp_str)$hypothesis

  # ---- Print
  cat("\n--- ", label, "  [", family_name, "] ---\n", sep = "")
  cat(sprintf("Link-scale coefficient:  %.4f   [95%% HDI: %.4f, %.4f]\n",
              link_hdi$y, link_hdi$ymin, link_hdi$ymax))

  if (family_name == "lognormal") {
    # With sum-coded predictor (-0.5/+0.5), beta is exactly log(y_big / y_small).
    # So exp(beta) is the ratio big/small, and (exp(beta) - 1) * 100 is the % change.
    pct      <- (exp(beta) - 1) * 100
    pct_hdi  <- median_hdi(pct, .width = 0.95)
    ratio_hdi <- median_hdi(exp(beta), .width = 0.95)
    cat(sprintf("Multiplicative effect:   %.3f×  [95%% HDI: %.3f×, %.3f×]\n",
                ratio_hdi$y, ratio_hdi$ymin, ratio_hdi$ymax))
    cat(sprintf("As %% change (big - small): %.2f%% [95%% HDI: %.2f%%, %.2f%%]\n",
                pct_hdi$y, pct_hdi$ymin, pct_hdi$ymax))
  } else if (family_name == "bernoulli") {
    or     <- exp(beta)
    or_hdi <- median_hdi(or, .width = 0.95)
    cat(sprintf("Odds ratio:              %.3f   [95%% HDI: %.3f, %.3f]\n",
                or_hdi$y, or_hdi$ymin, or_hdi$ymax))
  }
  # For gaussian the link IS the response scale, so nothing to back-transform.

  cat(sprintf("P(%s | data)  =  %.4f\n", hyp_str, hyp$Post.Prob))

  invisible(list(link = link_hdi, hypothesis = hyp))
}
```

### Directional hypothesis tests

For each pre-registered direction, we report the back-transformed effect (% change for
lognormal outcomes, odds ratio for Bernoulli) together with the posterior probability that the
coefficient is on the predicted side of zero. These are the numbers that anchor the prose
summary further down.

```{r report-effects-primary}
# Size effects (predicted direction)
report_effect(m_vdur,     "size_c",                direction = "positive",
              label = "Size effect on vowel duration (production)")
report_effect(m_f0,       "size_c",                direction = "negative",
              label = "Size effect on f0 (production)")
report_effect(m_perc_dur, "vowel_dur_ms_mean_z",   direction = "positive",
              label = "Vowel duration → big judgment (perception)")
report_effect(m_perc_f0,  "f0_mean_z",             direction = "negative",
              label = "f0 → big judgment (perception)")

# Animacy effects: modulation and baseline shift (direction open)
report_effect(m_vdur, "size_c:animacy_c", direction = "auto",
              label = "Animacy modulation of vowel duration")
report_effect(m_f0,   "size_c:animacy_c", direction = "auto",
              label = "Animacy modulation of f0")

report_effect(m_vdur,     "animacy_c", direction = "auto",
              label = "Animacy main effect on vowel duration")
report_effect(m_f0,       "animacy_c", direction = "auto",
              label = "Animacy main effect on f0")
report_effect(m_perc_dur, "animacy_c", direction = "auto",
              label = "Animacy main effect on 'big' judgment (duration model)")
report_effect(m_perc_f0,  "animacy_c", direction = "auto",
              label = "Animacy main effect on 'big' judgment (f0 model)")
```

### Posterior diagnostics for all four primary models

A walk through each of the four primary models in turn: posterior tables, PPCs, and half-eye
plots of all fixed effects. This is the place a sceptical reader can sanity-check the model
fits.

```{r report-primary}
walk2(
  list(m_vdur, m_f0, m_perc_dur, m_perc_f0),
  c("Production: vowel duration",
    "Production: f0",
    "Perception: duration → big judgment",
    "Perception: f0 → big judgment"),
  ~ report_model(.x, label = .y)
)
```

### Summary of primary findings

**Production – size effects.**
Speakers produced big referents with vowels **~18% longer** than small referents
(β = 0.16 on the log scale, 95% HDI [0.02, 0.30]; P(β > 0) = 0.99) and with
**~17% lower f0** (β = −0.18, 95% HDI [−0.28, −0.08]; P(β < 0) = 1.00).
Both predicted production effects are supported.

**Perception – f0 yes, duration no.**
Listeners used **f0 as a cue to size**: lower f0 increased the odds of a "big"
judgment by a factor of ~0.60 per +1 SD of speaker f0 (β = −0.52, 95% HDI
[−0.77, −0.25]; OR = 0.60; P(β < 0) ≈ 1.00). The f0 perception prediction is
supported. Listeners did **not** show a comparable effect for vowel duration
(β = 0.05, 95% HDI [−0.14, 0.25]; OR = 1.05; P(β > 0) = 0.70); the duration
perception prediction is not supported. The frequency-code channel is therefore
bidirectional in this dataset (produced and perceived); the duration channel is
asymmetric – speakers produce iconic lengthening for big referents, but listeners
do not appear to use vowel duration as a cue to size.

**Animacy as a modulator of the size effect – no evidence.**
The size × animacy interaction was close to zero on both acoustic dimensions
(vowel duration: β = 0.04, 95% HDI [−0.22, 0.30], P = 0.61; f0: β = −0.11,
95% HDI [−0.31, 0.08], P = 0.87). The size effect did not differ measurably
between animate and inanimate referents – iconic size encoding operates the
same way across both categories.

**Animacy as an independent shift in the baseline – credible in production, but see §7.8**
Animacy lands on the main-effect channel rather than the interaction channel.
Independent of size, animate referents were produced with vowels **~14% shorter**
than inanimate referents (β = −0.15, 95% HDI [−0.28, −0.02]; P(β < 0) = 0.99) and
with **~11% higher f0** (β = 0.10, 95% HDI [0.00, 0.20]; P(β > 0) = 0.98).
On the perception side both animacy main effects pointed in the same direction
(animate referents less likely to be judged "big") but their credible intervals
crossed zero (duration model: β = −0.92, 95% HDI [−2.18, 0.36], P = 0.93;
f0 model: β = −0.71, 95% HDI [−1.78, 0.39], P = 0.91). §7.8 returns to these
production animacy main effects and shows that they do not survive the addition of
an item-level cuteness covariate – the animacy effect shrinks toward zero and
cuteness itself does not take its place, consistent with the two predictors being
too correlated in this stimulus set to identify separately.

**Reading the main-effect-without-interaction pattern.** That speakers shift
their acoustic baseline by animacy *without* changing how strongly they encode
size means two things happen in parallel: (i) the size mapping is robust across
the animacy distinction, and (ii) there is a separate, additive shift between
speaking for animals and speaking for objects. Whether this shift reflects
animacy as a feature or systematic differences between the animal and object
item sets in this stimulus battery cannot be separated with the present design –
§7.8 makes this concrete by adding a separately-collected per-item cuteness
covariate, with the result that the animacy main effect on production becomes
non-credible without cuteness itself replacing it.

## 6.7 Size effects across the four primary models

A single forest plot of the size posterior across the four primary models – the headline graphic
for the size question. Positive values indicate the size-iconic direction for duration (longer
for big); for f0 the prediction is **negative** (lower for big), so a posterior concentrated
below zero supports the frequency-code prediction [@ohala_ethological_1984; @winter_size_2025].

```{r forest-main, fig.height=3.5, dpi=600, fig.retina=1, fig.cap=capFig("Size effect across the four primary models. Filled intervals: 66% and 95% HDI; dashed line at zero.")}
extract_term <- function(model, term, name) {
  as_draws_df(model) |>
    transmute(model = name, value = .data[[paste0("b_", term)]])
}

main_effects <- bind_rows(
  extract_term(m_vdur,     "size_c",                "vowel duration (production, log)"),
  extract_term(m_f0,       "size_c",                "f0 (production, log)"),
  extract_term(m_perc_dur, "vowel_dur_ms_mean_z",   "duration → big (perception, log-odds)"),
  extract_term(m_perc_f0,  "f0_mean_z",             "f0 → big (perception, log-odds)")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(main_effects, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "Effect estimate (big vs. small, or per +1 SD)",
       y = NULL#,
       #title = "Size effects in production and perception",
       #subtitle = "Predicted direction: positive for duration; negative for f0"
       )
```

```{r speaker-slopes, fig.width=7, fig.height=8, fig.cap=capFig("Per-speaker estimated size effect on vowel duration (log scale), sorted by magnitude. Dashed line = population-level estimate.")}
speaker_slopes_vdur <- m_vdur |>
  spread_draws(b_size_c, r_speaker[speaker, term]) |>
  filter(term == "size_c") |>
  mutate(slope = b_size_c + r_speaker) |>
  group_by(speaker) |>
  median_hdi(slope, .width = 0.95) |>
  ungroup() |>
  mutate(speaker = fct_reorder(factor(speaker), slope))

pop_med <- median(as_draws_df(m_vdur)$b_size_c)

ggplot(speaker_slopes_vdur, aes(y = speaker, x = slope, xmin = .lower, xmax = .upper)) +
  geom_pointrange(size = 0.3) +
  geom_vline(xintercept = 0,       linetype = 3, colour = "grey50") +
  geom_vline(xintercept = pop_med, linetype = 2, colour = "steelblue") +
  labs(x = "Per-speaker size effect on vowel duration (log)",
       y = NULL,
       title = "Speaker-level iconicity: how much each speaker stretches vowels for big",
       subtitle = "Dashed line = population estimate; dotted line = no effect")
```

```{r item-judgments, fig.width=7, fig.height=8, fig.cap=capFig("Posterior probability of 'big' judgment for each item (m_perc_f0), holding f0 at its mean. Items are coloured by their portrayed size.")}
item_size <- perc |> distinct(item_english, size)

item_probs <- m_perc_f0 |>
  spread_draws(b_Intercept, r_item_english[item_english, term]) |>
  filter(term == "Intercept") |>
  mutate(p_big = plogis(b_Intercept + r_item_english)) |>
  group_by(item_english) |>
  median_hdi(p_big, .width = 0.95) |>
  ungroup() |>
  left_join(item_size, by = "item_english") |>
  mutate(item_english = fct_reorder(item_english, p_big))

ggplot(item_probs, aes(y = item_english, x = p_big,
                       xmin = .lower, xmax = .upper, colour = size)) +
  geom_pointrange(size = 0.3) +
  geom_vline(xintercept = 0.5, linetype = 2, colour = "grey50") +
  scale_colour_manual(values = colorBlindBlack8[c(3, 4)]) +
  labs(x = "P(\"big\" judgment) at mean acoustic features",
       y = NULL,
       title = "Item-level baseline 'big-judgability' (perception)",
       subtitle = "Listeners mostly identify the referent from the picture; acoustic cues modulate on top of this")
```

## 6.8 Animacy effects across the four primary models

Each model contains two animacy-relevant terms: the **interaction** `size_c × animacy_c`
(does the size effect differ between animate and inanimate referents?) and the **main
effect** of `animacy_c` (does animacy shift the outcome independent of size?). A finding can
land on either, both, or neither. Reporting both keeps us honest about what the data say,
since the animacy expectation was open as to which channel would carry it.

```{r forest-interaction, fig.height=3.5, fig.cap=capFig("Size × animacy interaction across the four primary models.")}
interactions_df <- bind_rows(
  extract_term(m_vdur,     "size_c:animacy_c",                  "vowel duration (production)"),
  extract_term(m_f0,       "size_c:animacy_c",                  "f0 (production)"),
  extract_term(m_perc_dur, "vowel_dur_ms_mean_z:animacy_c",     "duration (perception)"),
  extract_term(m_perc_f0,  "f0_mean_z:animacy_c",               "f0 (perception)")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(interactions_df, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "Size × Animacy interaction estimate",
       y = NULL,
       title = "Does animacy modulate the size effect?",
       subtitle = "Nonzero = size effect differs between animate and inanimate referents")
```

```{r forest-animacy-main, fig.height=3.5, fig.cap=capFig("Animacy main effect across the four primary models.")}
animacy_main_df <- bind_rows(
  extract_term(m_vdur,     "animacy_c", "vowel duration (production)"),
  extract_term(m_f0,       "animacy_c", "f0 (production)"),
  extract_term(m_perc_dur, "animacy_c", "duration model (perception)"),
  extract_term(m_perc_f0,  "animacy_c", "f0 model (perception)")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(animacy_main_df, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "Animacy main effect estimate",
       y = NULL,
       title = "Does animacy shift the outcome independent of size?",
       subtitle = "Nonzero = animate and inanimate referents have different baselines")
```

```{r parallel-lines-helper}
# Predicted outcome at the four corners of size × animacy.
# Random effects are integrated out (re_formula = NA): we get population-level
# predictions, which is what makes the parallel-vs-tilted geometry clean.
plot_size_by_animacy <- function(model, y_label, title_text,
                                 vowel_ref = "a:", phoneme_ref = "C", sex_ref = "m") {
  # Use the model's actual data columns to decide which covariates to supply.
  # all.vars(formula(model)) doesn't reliably traverse a brmsformula tree.
  model_vars <- names(model$data)

  newd <- expand_grid(
    size_c    = c(-0.5, 0.5),
    animacy_c = c(-0.5, 0.5)
  )
  if ("vowel"   %in% model_vars) newd$vowel   <- vowel_ref
  if ("phoneme" %in% model_vars) newd$phoneme <- phoneme_ref
  if ("sex"     %in% model_vars) newd$sex     <- sex_ref

  newd |>
    add_epred_draws(model, re_formula = NA) |>
    mutate(
      size    = factor(if_else(size_c < 0, "small", "big"), levels = c("small", "big")),
      animacy = if_else(animacy_c < 0, "inanimate", "animate")
    ) |>
    ggplot(aes(x = size, y = .epred, colour = animacy, group = animacy)) +
    stat_pointinterval(.width = c(0.66, 0.95),
                       position = position_dodge(width = 0.2)) +
    stat_summary(fun = median, geom = "line",
                 position = position_dodge(width = 0.2),
                 linewidth = 0.8) +
    scale_colour_manual(values = colorBlindBlack8[c(3, 4)]) +
    labs(x = NULL, y = y_label, colour = "Animacy", title = title_text)
}
```

```{r size-by-animacy-plots, fig.width=10, fig.height=4.5, fig.cap=capFig("Population-level predicted outcomes at the four cells of size × animacy. Parallel lines indicate an animacy main effect without modulation; tilted (non-parallel) lines would indicate that the size effect itself differs between animate and inanimate referents.")}
p1 <- plot_size_by_animacy(m_vdur, "Vowel duration (ms)", "Vowel duration")
p2 <- plot_size_by_animacy(m_f0,   "f0 (Hz)",             "f0")

(p1 + p2) + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

Look at the *vowel duration* panel: green (inanimate) sits above blue (animate), both climb at essentially the same rate from small → big. That's the textbook parallel-lines pattern – animacy shifts the baseline (the ~14% gap), size effect is the same for both (the ~17% lift). This is the `m_vdur` model: a credible animacy main effect at this point in the analysis, with the interaction at zero. §7.8 will revisit how stable this main effect is when item-level cuteness is added to the model.

The f0 panel tells a slightly different story. The lines aren't quite parallel – blue (animate) starts higher at `small` (~**148** Hz) and converges with green (inanimate) at `big` (~**115** Hz). That mild convergence is exactly the soft hint of a non-zero interaction we saw in the numbers (β = −0.11, 95% **HDI** [−0.31, 0.08]; P(β < 0) = 0.87). The CrI brushes zero so it's not a claim, but the geometry makes the hint visible: animate referents seem to do more iconic "lowering" than inanimate ones do.

# 7. Exploratory analyses

These models reuse the primary structure with swapped outcomes or predictors. They are deliberately exploratory: we are examining the broader acoustic signal for places where size iconicity (or animacy) may leak into channels that have not been examined before, in the spirit of the pluripotential view of iconicity. No directional hypotheses are committed to here, and no claim of robustness is made pending replication. The §7.7 takeaways below distil what we saw; the §8 forests give the visual synthesis.

## 7.1 Templates

Three reusable fitters. `fit_prod_vocalic` mirrors the production frame in §6.3 (size × animacy interaction, sex covariate, vowel fixed effect, by-speaker random intercept and slope, by-item random intercept). `fit_prod_fric` is the same idea for the fricative tables, with `phoneme` replacing `sex + vowel` as the covariate. `fit_perc_bin` mirrors §6.4–6.5 for a binary perception outcome on a single z-scored acoustic predictor.

```{r exploratory-template}
# Production: continuous vocalic outcome
fit_prod_vocalic <- function(outcome, family = gaussian(), data = vowel_long) {
  formula <- bf(as.formula(paste(
    outcome, "~ size_c * animacy_c + sex + vowel +",
    "(1 + size_c | speaker) + (1 | item_english)"
  )))
  d <- data |> filter(!is.na(.data[[outcome]]))
  fit_brm(formula, data = d, family = family,
          file = paste0(models, "m_prod_", outcome))
}

# Production: continuous fricative outcome
fit_prod_fric <- function(outcome, family = gaussian(), data = fric_long) {
  formula <- bf(as.formula(paste(
    outcome, "~ size_c * animacy_c + phoneme +",
    "(1 + size_c | speaker) + (1 | item_english)"
  )))
  d <- data |> filter(!is.na(.data[[outcome]]))
  fit_brm(formula, data = d, family = family,
          file = paste0(models, "m_prod_", outcome))
}

# Perception: binary outcome on a single z-scored predictor
fit_perc_bin <- function(outcome, predictor_z, data = perc) {
  formula <- bf(as.formula(paste(
    outcome, "~", predictor_z, "* animacy_c +",
    "(1 | listener) + (1 | speaker) + (1 | item_english)"
  )))
  d <- data |> filter(!is.na(.data[[predictor_z]]))
  fit_brm(formula, data = d, family = bernoulli(),
          file = paste0(models, "m_perc_", outcome, "_", predictor_z))
}
```

## 7.2 Other vocalic features (F1, F2, CPP, H1–H2)

We model the same vocalic frame as in §6 but swap the outcome. F1 and F2 are the first two formants (correlates of vowel quality). F1 in particular tracks both mouth opening and tongue height: lower F1 reflects a less open mouth or a higher tongue position, and the two articulators trade off but not one-to-one. CPP (cepstral peak prominence) and H1–H2 (the level difference between the first and second harmonics) index voice quality: higher CPP indicates a more periodic, modal voice; H1–H2 is typically larger in breathier voice. F1 and F2 are modelled lognormally because they are bounded below at zero with a right tail; CPP and H1–H2 are modelled with the default Gaussian since they are centred dB-scale quantities.

The formant question is worth flagging up front. @cwiek_scope_2025 examined whether object size modulates lip opening or formant dispersion in a paradigm where participants "shot" at cans of different sizes mounted on a wall while saying *piff* or *paff*, and reported no detectable size effect on either measure: vowel identity dominated the articulatory and acoustic outcomes. Our paradigm differs (imitative production of the same carrier sentence with free choice of vocalic realisation rather than a fixed two-vowel target, and animate referents alongside inanimate ones), so any divergence between the two studies is more usefully read as a difference in scope than as a contradiction. We examine F1 and F2 here in that spirit.

```{r exploratory-vocalic, results='hide'}
m_F1   <- fit_prod_vocalic("F1",   family = lognormal())
m_F2   <- fit_prod_vocalic("F2",   family = lognormal())
m_CPP  <- fit_prod_vocalic("CPP")          # gaussian (default)
m_H1H2 <- fit_prod_vocalic("H1H2")         # gaussian (default)
```

```{r expl-vocalic-display-F1, fig.height=3, fig.cap=capFig("Posterior predictive check for F1 (lognormal).")}
summary(m_F1)
pp_check(m_F1, ndraws = 100) + ggtitle("PPC – F1")
```

Bimodal because of different vowels.

```{r expl-vocalic-display-F2, fig.height=3, fig.cap=capFig("Posterior predictive check for F2 (lognormal).")}
summary(m_F2)
pp_check(m_F2, ndraws = 100) + ggtitle("PPC – F2")
```

Bimodal because of different vowels.

```{r expl-vocalic-display-CPP, fig.height=3, fig.cap=capFig("Posterior predictive check for CPP (Gaussian).")}
summary(m_CPP)
pp_check(m_CPP, ndraws = 100) + ggtitle("PPC – CPP")
```

```{r expl-vocalic-display-H1H2, fig.height=3, fig.cap=capFig("Posterior predictive check for H1–H2 (Gaussian).")}
summary(m_H1H2)
pp_check(m_H1H2, ndraws = 100) + ggtitle("PPC – H1–H2")
```

## 7.3 Fricative features (production)

Fricative spectral cues are routinely linked to perceived size and to the geometry of the vocal tract [@jongman_acoustic_2000; @al-tamimi_jalalal-tamimipraat-spectral-analyses-fricatives-stops_2022]. The four features we model here are fricative duration (ms), fricative SPL (a dB-scale intensity measure on the whole fricative segment), peak spectral frequency (Hz), and centre of gravity (Hz). Duration is lognormal-distributed; the other three are kept on their native scales with a Gaussian likelihood (taking the log of frequency values tends to overshoot the actual right-skew).

```{r exploratory-fric, results='hide'}
m_fdur  <- fit_prod_fric("fric_dur_ms",  family = lognormal())
m_fSPL  <- fit_prod_fric("fric_SPL_dB")              # gaussian
m_fpeak <- fit_prod_fric("fric_peak_Hz")             # gaussian (log overshoots)
m_fCoG  <- fit_prod_fric("fric_CoG_Hz")              # gaussian (log overshoots)
```

```{r expl-fric-display-fdur, fig.height=3, fig.cap=capFig("Posterior predictive check for fricative duration (lognormal).")}
summary(m_fdur)
pp_check(m_fdur, ndraws = 100) + ggtitle("PPC – fricative duration")
```

```{r expl-fric-display-fSPL, fig.height=3, fig.cap=capFig("Posterior predictive check for fricative SPL (Gaussian).")}
summary(m_fSPL)
pp_check(m_fSPL, ndraws = 100) + ggtitle("PPC – fricative SPL")
```

```{r expl-fric-display-fpeak, fig.height=3, fig.cap=capFig("Posterior predictive check for fricative peak frequency (Gaussian).")}
summary(m_fpeak)
pp_check(m_fpeak, ndraws = 100) + ggtitle("PPC – fricative peak frequency")
```

```{r expl-fric-display-fCoG, fig.height=3, fig.cap=capFig("Posterior predictive check for fricative centre of gravity (Gaussian).")}
summary(m_fCoG)
pp_check(m_fCoG, ndraws = 100) + ggtitle("PPC – fricative centre of gravity")
```

## 7.4 Perception of exploratory features

These models mirror §6.4–6.5 but swap the acoustic predictor. Each tests whether listeners use the speaker's exploratory feature (z-scored within speaker) to call "big". As before the interaction with `animacy_c` is included so the perception models can answer both whether the cue matters on average *and* whether animacy modulates its use.

```{r exploratory-perc, results='hide'}
m_perc_F1     <- fit_perc_bin("big_judgment", "F1_mean_z")
m_perc_F2     <- fit_perc_bin("big_judgment", "F2_mean_z")
m_perc_CPP   <- fit_perc_bin("big_judgment", "CPP_mean_z")
m_perc_H1H2  <- fit_perc_bin("big_judgment", "H1H2_mean_z")
m_perc_fCoG  <- fit_perc_bin("big_judgment", "fric_CoG_Hz_mean_z")
m_perc_fdur   <- fit_perc_bin("big_judgment", "fric_dur_ms_mean_z")
m_perc_fSPL   <- fit_perc_bin("big_judgment", "fric_SPL_dB_mean_z")
m_perc_fpeak  <- fit_perc_bin("big_judgment", "fric_peak_Hz_mean_z")
```

```{r expl-perc-display-F1, fig.height=3, fig.cap=capFig("Perception: F1 as cue to 'big' judgment.")}
summary(m_perc_F1)
pp_check(m_perc_F1, ndraws = 100) + ggtitle("PPC – perception of F1")
```

```{r expl-perc-display-F2, fig.height=3, fig.cap=capFig("Perception: F2 as cue to 'big' judgment.")}
summary(m_perc_F2)
pp_check(m_perc_F2, ndraws = 100) + ggtitle("PPC – perception of F2")
```

```{r expl-perc-display-CPP, fig.height=3, fig.cap=capFig("Perception: CPP as cue to 'big' judgment.")}
summary(m_perc_CPP)
pp_check(m_perc_CPP, ndraws = 100) + ggtitle("PPC – perception of CPP")
```

```{r expl-perc-display-H1H2, fig.height=3, fig.cap=capFig("Perception: H1H2 as cue to 'big' judgment.")}
summary(m_perc_H1H2)
pp_check(m_perc_H1H2, ndraws = 100) + ggtitle("PPC – perception of H1H2")
```

```{r expl-perc-display-fCoG, fig.height=3, fig.cap=capFig("Perception: fCoG as cue to 'big' judgment.")}
summary(m_perc_fCoG)
pp_check(m_perc_fCoG, ndraws = 100) + ggtitle("PPC – perception of fCoG")
```

```{r expl-perc-display-fdur, fig.height=3, fig.cap=capFig("Perception: fricative duration as cue to 'big' judgment.")}
summary(m_perc_fdur)
pp_check(m_perc_fdur, ndraws = 100) + ggtitle("PPC – perception of fricative duration")
```

```{r expl-perc-display-fSPL, fig.height=3, fig.cap=capFig("Perception: fricative SPL as cue to 'big' judgment.")}
summary(m_perc_fSPL)
pp_check(m_perc_fSPL, ndraws = 100) + ggtitle("PPC – perception of fricative SPL")
```

```{r expl-perc-display-fpeak, fig.height=3, fig.cap=capFig("Perception: fricative peak frequency as cue to 'big' judgment.")}
summary(m_perc_fpeak)
pp_check(m_perc_fpeak, ndraws = 100) + ggtitle("PPC – perception of fricative peak frequency")
```

## 7.5 Does the *animacy* judgment depend on acoustic cues?

These two models flip the perception outcome from "big" to "animate" and ask whether listeners draw on vowel duration or f0 to decide whether they are hearing an animal. Treating animacy as an outcome here is sound-symbolism in the broader sense: the same acoustic dimensions that encode size might also leak into the animate/inanimate distinction.

```{r rq4, results='hide'}
m_anim_dur <- fit_perc_bin("animate_judgment", "vowel_dur_ms_mean_z")
m_anim_f0  <- fit_perc_bin("animate_judgment", "f0_mean_z")
```

```{r expl-anim-display-dur, fig.height=3, fig.cap=capFig("Perception: vowel duration as cue to animacy judgment.")}
summary(m_anim_dur)
pp_check(m_anim_dur, ndraws = 100) + ggtitle("PPC – vowel duration → animate judgment")
```

```{r expl-anim-display-f0, fig.height=3, fig.cap=capFig("Perception: f0 as cue to animacy judgment.")}
summary(m_anim_f0)
pp_check(m_anim_f0, ndraws = 100) + ggtitle("PPC – f0 → animate judgment")
```

## 7.6 A note on multinomial category judgments

A logical extension would be to model the full 6-way `picked_category` outcome (bird, mammal, reptile, appliance, vehicle, instrument) as a multinomial response to the same acoustic cues. We considered this and decided against including it in the main analysis for two reasons. First, a categorical 6-level outcome multiplies the coefficient space (5 contrasts × every fixed effect) and is hard to summarize compactly in a short-format paper. Second, it does not speak directly to our expectations: the size question is covered by §6, and the animacy question is covered by §6 (for animacy as production-side baseline) and §7.5 (for animacy as a perceptual outcome on its own). The chunk below is left in the script for transparency – it can be turned on by setting `eval=TRUE` – but is not part of the reported analysis. The model also does not add a parallel f0 version because the animacy-judgment perception model in §7.5 already captures the relevant question for that channel.

```{r rq3, eval=FALSE, results='hide'}
m_cat_dur <- fit_brm(
  bf(picked_category ~ vowel_dur_ms_mean_z * animacy_c +
       (1 | listener) + (1 | speaker) + (1 | item_english)),
  data   = perc |> filter(!is.na(vowel_dur_ms_mean_z)) |>
                   mutate(picked_category = factor(picked_category)),
  family = categorical(),
  file   = paste0(models, "m_cat_dur")
)
```

## 7.7 Exploratory takeaways

These results are read as descriptive flags, not robust findings, and we mark the directions
that look worth following up.

**Vocalic features beyond duration and f0.** F1 is credibly lower for big referents on the log scale (β = -0.06, 95% HDI [-0.10, -0.03]; ~6% lower F1 for big). Mechanistically a lower F1 can come from either a less open mouth or a higher tongue position (the two move together but not one-to-one), so the direction is at face value the *opposite* of the open-vowel-for-big
stereotype – our speakers are realizing big referents with a slightly more closed or
higher-tongue vocalic gesture, not a more open one. The direction also runs against
@cwiek_scope_2025's null on lip opening and formant dispersion for object size, but the
paradigms differ enough (their participants "shot" at cans of different sizes on the wall
while saying *piff* or *paff*; ours imitatively portrayed both animate and inanimate
referents in a carrier sentence) that we treat the contrast as hypothesis-generating rather
than as evidence against their conclusion. F2 shows no credible size effect. CPP is
credibly higher for big referents (β = 1.10 dB, 95% HDI [0.11, 2.11]),
consistent with more periodic voicing for big-referent productions, and credibly
lower for animate referents (animacy main effect β = -1.57, 95% HDI [-2.44, -0.69]) –
with the same caveat as in §7.8: the animacy main effect on any of these channels
may reflect item-level dimensions that travel with the animate/inanimate split
rather than animacy as a feature. H1–H2 is also credibly lower for big referents (β = −1.21 dB, 95% HDI [−2.31, −0.11]) – with CPP, a converging hint that big-referent productions carry a less breathy, more pressed voice.
On the perception side, listeners used neither CPP (β = −0.09, 95% HDI [−0.27, 0.09]) nor H1–H2 (β = 0.03, 95% HDI [−0.15, 0.21]) to judge size: the voice-quality channel, like vowel duration, is produced here but not decoded.

**Fricative production.** Fricative SPL is louder for big referents by roughly 3 dB
(β = 3.08, 95% HDI [0.61, 5.61]) – the clearest iconic effect among the fricative measures.
Fricative duration is in the size-iconic direction but its 95% HDI brushes zero (β = 0.10,
95% HDI [-0.02, 0.21]; P(β > 0) ≈ 0.95). Fricative peak frequency and centre of gravity show
no credible size effect; phoneme identity, as expected, dominates both.

**Fricative perception.** Listeners use fricative duration as a cue to size
(β = 0.25, 95% HDI [0.04, 0.45]) and the use is amplified for animate referents
(interaction β = 0.50, 95% HDI [0.09, 0.91]). Listeners also use fricative SPL
(β = 0.24, 95% HDI [0.05, 0.43]). F1, F2, fricative peak frequency, and fricative centre of gravity are not credibly used.
The fricative-duration result is interesting on its own: it is a perception-side effect for a
production-side feature whose production-side credibility was marginal – a pattern more often
seen the other way round in this dataset.

**Animacy judgment.** Listeners judge productions as animate more often when speaker f0 is
higher (β = 0.23 log-odds per +1 SD, 95% HDI [0.01, 0.45]; OR ≈ 1.26). The corresponding
duration model is null (β = -0.10, 95% HDI [-0.28, 0.09]). The interaction with portrayed
animacy in the f0 model is suggestive but its HDI crosses zero (β = 0.41, 95% HDI [-0.03, 0.87]).
The pattern is consistent with f0 carrying both a size signal (lower f0 → bigger) and an
animacy signal (higher f0 → animate), and the two are dissociable in the data.

## 7.8 Cuteness as a candidate explanation for the animacy main effect

The §6 finding that animate referents are produced with shorter vowels and
higher f0 *regardless of their size* is striking but interpretively
ambiguous: the animal stimuli differ from the object stimuli on more than
just animacy. One obvious candidate is **perceived cuteness**: a separate
group of 17 raters scored each stimulus picture for cuteness on a 1–7 scale.
Cuteness covaries with animacy (animals tend to be rated cuter than
appliances and vehicles) and partly with size (smaller things tend to be
cuter than bigger things), so it is a plausible carrier of what the §6
animacy main effect is picking up.

We test this in the same way the original animacy main effect was tested:
refit each of the four primary-claim models *plus* the animate-judgment
perception model from §7.5, adding `cuteness_z` (z-scored per-item mean
cuteness) as a fixed effect alongside `size_c * animacy_c`. The question
is descriptive: does the animacy coefficient shrink toward zero when
cuteness is partialled out, and does cuteness itself carry a credible
effect on each outcome?

The random-effects structure is unchanged. Cuteness varies only between
items, so it does not enter as a varying slope at the speaker or listener
level – the `(1 | item_english)` intercept still absorbs the
between-item residual after cuteness is accounted for.

### 7.8.1 Augmented production models

```{r m-cute-prod, results='hide'}
m_vdur_cute <- fit_brm(
  bf(vowel_dur_ms ~ size_c * animacy_c + cuteness_z + vowel +
       (1 + size_c | speaker) + (1 | item_english)),
  data   = vowel_long |> filter(!is.na(vowel_dur_ms), !is.na(cuteness_z)),
  family = lognormal(),
  file   = paste0(models, "m_vdur_cute")
)

m_f0_cute <- fit_brm(
  bf(f0 ~ size_c * animacy_c + cuteness_z + sex + vowel +
       (1 + size_c | speaker) + (1 | item_english)),
  data   = vowel_long |> filter(!is.na(f0), !is.na(cuteness_z)),
  family = lognormal(),
  file   = paste0(models, "m_f0_cute")
)
```

```{r m-cute-prod-display}
summary(m_vdur_cute)
summary(m_f0_cute)
```

### 7.8.2 Augmented perception models

```{r m-cute-perc, results='hide'}
m_perc_f0_cute <- fit_brm(
  bf(big_judgment ~ f0_mean_z * animacy_c + cuteness_z +
       (1 | listener) + (1 | speaker) + (1 | item_english)),
  data   = perc |> filter(!is.na(f0_mean_z), !is.na(cuteness_z)),
  family = bernoulli(),
  file   = paste0(models, "m_perc_f0_cute")
)

m_anim_f0_cute <- fit_brm(
  bf(animate_judgment ~ f0_mean_z * animacy_c + cuteness_z +
       (1 | listener) + (1 | speaker) + (1 | item_english)),
  data   = perc |> filter(!is.na(f0_mean_z), !is.na(cuteness_z)),
  family = bernoulli(),
  file   = paste0(models, "m_anim_f0_cute")
)
```

```{r m-cute-perc-display}
summary(m_perc_f0_cute)
summary(m_anim_f0_cute)
```

### 7.8.3 Directional tests on the cuteness coefficient

```{r m-cute-effects}
report_effect(m_vdur_cute,    "cuteness_z", direction = "auto",
              label = "Cuteness → vowel duration (production)")
report_effect(m_f0_cute,      "cuteness_z", direction = "auto",
              label = "Cuteness → f0 (production)")
report_effect(m_perc_f0_cute, "cuteness_z", direction = "auto",
              label = "Cuteness → 'big' judgment (perception, f0 model)")
report_effect(m_anim_f0_cute, "cuteness_z", direction = "auto",
              label = "Cuteness → 'animate' judgment (perception, f0 model)")

# And the animacy main effect in each augmented model – the comparison number
report_effect(m_vdur_cute,    "animacy_c", direction = "auto",
              label = "Animacy main effect on vowel duration AFTER controlling for cuteness")
report_effect(m_f0_cute,      "animacy_c", direction = "auto",
              label = "Animacy main effect on f0 AFTER controlling for cuteness")
report_effect(m_perc_f0_cute, "animacy_c", direction = "auto",
              label = "Animacy main effect on 'big' judgment AFTER controlling for cuteness")
report_effect(m_anim_f0_cute, "animacy_c", direction = "auto",
              label = "Animacy main effect on 'animate' judgment AFTER controlling for cuteness")
```

### 7.8.4 Side-by-side: animacy posterior with and without cuteness

This is the visual that answers the question. For each outcome, we overlay
the posterior of the animacy main-effect coefficient from the baseline
model (§§6.2–6.5, §7.5) and from the cuteness-augmented model (§§7.8.1–2).
If cuteness is doing the work the animacy main effect was credited for in
§6, the augmented posterior should shrink toward zero relative to the
baseline.

```{r animacy-shift, fig.height=5.5, dpi=600, fig.retina=1, fig.width=9, fig.cap=capFig("Animacy main-effect posterior in the baseline model vs. the cuteness-augmented model. Shrinkage toward zero indicates that cuteness absorbs the animacy effect.")}
extract_animacy <- function(model, outcome_name, model_kind) {
  as_draws_df(model) |>
    transmute(outcome    = outcome_name,
              model_kind = model_kind,
              value      = b_animacy_c)
}

animacy_compare <- bind_rows(
  extract_animacy(m_vdur,         "vowel duration (production, log)", "baseline"),
  extract_animacy(m_vdur_cute,    "vowel duration (production, log)", "+ cuteness"),
  extract_animacy(m_f0,           "f0 (production, log)",             "baseline"),
  extract_animacy(m_f0_cute,      "f0 (production, log)",             "+ cuteness"),
  extract_animacy(m_perc_f0,      "f0 → big judgment (logit)",        "baseline"),
  extract_animacy(m_perc_f0_cute, "f0 → big judgment (logit)",        "+ cuteness"),
  extract_animacy(m_anim_f0,      "f0 → animate judgment (logit)",    "baseline"),
  extract_animacy(m_anim_f0_cute, "f0 → animate judgment (logit)",    "+ cuteness")
) |>
  mutate(
    outcome    = factor(outcome,    levels = unique(outcome)),
    model_kind = factor(model_kind, levels = c("baseline", "+ cuteness"))
  )

ggplot(animacy_compare,
       aes(x = value, y = fct_rev(outcome), fill = model_kind)) +
  stat_halfeye(
    .width         = c(0.66, 0.95),
    alpha          = 0.85,
    slab_colour    = "white",       # fixed, not mapped — kills the 2nd legend
    slab_linewidth = 0.3,
    position       = position_dodge(width = 0.65)
  ) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  scale_fill_manual(values = colorBlindBlack8[c(3, 7)]) +
  labs(
    x        = "Animacy main-effect estimate (link scale)",
    y        = NULL,
    fill     = NULL#,
    #title    = "Does cuteness absorb the animacy main effect?",
    #subtitle = "Each outcome: baseline model (§6 / §7.5) vs. with cuteness_z added"
  ) +
  theme(
    plot.title          = element_text(size = 13, face = "bold"),
    plot.subtitle       = element_text(size = 10, colour = "grey30"),
    plot.title.position = "plot",          # lets title use the full plot width
    legend.position     = "top",
    legend.justification = "left",
    axis.text.y         = element_text(size = 10.5),
    axis.title.x        = element_text(size = 10.5),
    panel.grid.minor    = element_blank()
  )
```

### 7.8.5 Cuteness posterior across the five augmented models

```{r cute-forest, fig.height=3.5, fig.cap=capFig("Cuteness coefficient across the five augmented models. Production outcomes are on the log scale; perception outcomes are on the logit scale.")}
cute_forest <- bind_rows(
  extract_term(m_vdur_cute,    "cuteness_z", "vowel duration (production, log)"),
  extract_term(m_f0_cute,      "cuteness_z", "f0 (production, log)"),
  extract_term(m_perc_f0_cute, "cuteness_z", "f0 → big judgment (logit)"),
  extract_term(m_anim_f0_cute, "cuteness_z", "f0 → animate judgment (logit)")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(cute_forest, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "Cuteness coefficient (per +1 SD item cuteness)",
       y = NULL,
       title = "Effect of perceived cuteness across outcomes")
```

### 7.8.6 Reading

**Production: animacy shrinks, but cuteness doesn't replace it.** The
animacy main effect on both production outcomes shrinks toward zero when
cuteness enters the model. On vowel duration, the baseline coefficient
(β = −0.15, 95% HDI [−0.28, −0.02]) becomes β = −0.07 [−0.38, 0.22] –
the point estimate halves and the HDI widens to straddle zero. On f0,
the baseline coefficient (β = +0.10, 95% HDI [0.00, 0.20]) becomes
β = +0.02 [−0.19, 0.24]. But the cuteness coefficient does *not* take
over the variance the animacy effect lost: cuteness on duration is
β = −0.04 [−0.20, 0.11] (P(β < 0) = 0.71) and cuteness on f0 is
β = +0.05 [−0.07, 0.17] (P(β > 0) = 0.79). Neither predictor reaches
credibility in the augmented models.

The cleanest read is that animacy and cuteness are too tightly correlated
in this stimulus set (§3.4 already shows that animals are systematically
cuter than objects) to be separately identified from 36 items' worth of
between-item variance. The §6 animacy main effect on production is
therefore fragile under any item-level covariate that tracks the
animal-vs-object distinction; we cannot tell from the present data
whether the carrier is animacy *qua* feature, cuteness, or a third
item-level dimension confounded with both.

**Perception: a different pattern – suppressor-style instability rather
than absorption.** The big-judgment model behaves unusually when cuteness
enters. The baseline animacy coefficient (β = −0.92, 95% HDI
[−2.18, 0.36], P = 0.93) flips sign and inflates to β = +2.28
[0.68, 3.99], P = 0.998, while the cuteness coefficient becomes large
and negative (β = −1.83 [−2.70, −0.98]; OR = 0.16; P > 0.999) and the f0
effect itself is essentially unchanged (β = −0.52 [−0.77, −0.26]). An
animacy posterior that grows *and* flips sign while a correlated covariate
enters with a large opposite-sign coefficient is a classical suppressor
pattern: the two predictors are jointly identifying a contrast that
neither captures alone, and the individual coefficients on that contrast
cannot be interpreted as effects of animacy or of cuteness on their own.
We do not read a substantive claim off this fit.

The animate-judgment model is, by contrast, almost untouched by adding
cuteness: animacy stays at β ≈ +2.2 in both baseline and augmented fits,
and cuteness is null (β = −0.09 [−0.71, 0.49]). This is the expected
pattern when the outcome and a fixed effect index nearly the same thing –
listeners judge animals as animate from the picture, and there is
correspondingly little residual room for an item-level cuteness covariate
to add explanatory value.

**The §6 size effects are unaffected.** A useful side check from the
augmented production models: the size effect on vowel duration
(β = +0.14 [−0.03, 0.30]) and on f0 (β = −0.16 [−0.28, −0.04]) barely
move from their §6 values (+0.16 and −0.18 respectively), and the
perception-of-f0 effect (β = −0.52) is unchanged. The primary §6 claims
– that speakers reliably lengthen vowels and lower f0 for big referents,
and that listeners use f0 to call "big" – are therefore not contaminated
by item-level cuteness or animacy. Size iconicity rides on top of these
between-item dimensions, not through them.

**Bottom line for the §6 animacy story.** The cuteness analysis does not
cleanly support the §6 reading of an animacy main effect, but it also
does not cleanly support a "cuteness in disguise" alternative. What it
does support is the limitation already flagged in §9: in this stimulus
set, animacy is confounded with item-level dimensions (cuteness being
one we can name) to a degree that the present design cannot resolve.
The honest write-up is to keep §6 as a descriptive observation –
animate referents in this set were produced with shorter vowels and
higher f0, and listener big-judgments tracked f0 – without claiming
animacy *qua* feature as the carrier of the production-side baseline
shift. A design that orthogonalises animacy and cuteness, with more
items per cell, would be needed to identify either independently.

# 8. Cross-feature comparison

The three forest plots below give the visual synthesis of §§6–7. The first asks where size leaks
into the produced signal across all ten acoustic features we measured. The second asks which of
those features listeners use to judge size. The third looks at the same question for animacy
judgments instead of size judgments. Each panel is on its own scale (log multiplicative effects,
dB, Hz, or log-odds) – the comparisons within a panel are meaningful; the comparisons across
panels are not.

## 8.1 Production: where does size leak into the signal?

```{r forest-prod, fig.height=9, dpi=600, fig.retina=1, fig.cap=capFig("Production: size effect across all acoustic features, faceted by effect-size scale (log, dB, Hz). Each panel has its own x-axis range – magnitudes across panels are NOT directly comparable.")}
forest_prod <- bind_rows(
  extract_term(m_vdur,  "size_c", "vowel duration (log)"),
  extract_term(m_f0,    "size_c", "f0 (log)"),
  extract_term(m_F1,    "size_c", "F1 (log)"),
  extract_term(m_F2,    "size_c", "F2 (log)"),
  extract_term(m_fdur,  "size_c", "fricative duration (log)"),
  extract_term(m_CPP,   "size_c", "CPP (dB)"),
  extract_term(m_H1H2,  "size_c", "H1–H2 (dB)"),
  extract_term(m_fSPL,  "size_c", "fricative SPL (dB)"),
  extract_term(m_fpeak, "size_c", "fricative peak (Hz)"),
  extract_term(m_fCoG,  "size_c", "fricative CoG (Hz)")
) |>
  mutate(
    scale_group = case_when(
      str_detect(model, "\\(log\\)") ~ "Log scale (lognormal models)",
      str_detect(model, "\\(dB\\)")  ~ "dB scale (Gaussian)",
      str_detect(model, "\\(Hz\\)")  ~ "Hz scale (Gaussian)"
    ),
    scale_group = factor(scale_group,
                         levels = c("Log scale (lognormal models)",
                                    "dB scale (Gaussian)",
                                    "Hz scale (Gaussian)")),
    model = factor(model, levels = unique(model))
  )

ggplot(forest_prod, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  facet_wrap(~ scale_group, scales = "free", ncol = 1) +
  labs(x = "Effect of big vs. small (sum-coded)",
       y = NULL#,
       #title = "Production: size effect across all acoustic features",
       #subtitle = "Top panel includes the primary-claim outcomes (vowel duration, f0)"
       )
```
 
## 8.2 Perception: which cues drive "big" judgments?

```{r forest-perc, fig.height=4, dpi=600, fig.retina=1, fig.cap=capFig("Perception: log-odds shift in 'big' judgment per +1 SD of each acoustic feature. Top two rows = primary claim; below = exploratory.")}
forest_perc <- bind_rows(
  extract_term(m_perc_dur,   "vowel_dur_ms_mean_z", "vowel duration"),
  extract_term(m_perc_f0,    "f0_mean_z",           "f0"),
  extract_term(m_perc_F1,    "F1_mean_z",           "F1"),
  extract_term(m_perc_F2,    "F2_mean_z",           "F2"),
  extract_term(m_perc_CPP,   "CPP_mean_z",          "CPP"),
  extract_term(m_perc_H1H2,  "H1H2_mean_z",         "H1–H2"),
  extract_term(m_perc_fCoG,  "fric_CoG_Hz_mean_z",  "fricative CoG"),
  extract_term(m_perc_fdur,  "fric_dur_ms_mean_z",  "fricative duration"),
  extract_term(m_perc_fSPL,  "fric_SPL_dB_mean_z",  "fricative SPL"),
  extract_term(m_perc_fpeak, "fric_peak_Hz_mean_z", "fricative peak")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(forest_perc, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "log-odds shift of \"big\" judgment per +1 SD acoustic feature",
       y = NULL#,
       #title = "Perception: which cues drive \"big\" judgments?",
       #subtitle = "Top two rows = primary claim; below = exploratory"
       )
```

## 8.3 Perception: which cues drive "animate" judgments?

```{r forest-anim, fig.height=2.5, fig.cap=capFig("Perception: log-odds shift in 'animate' judgment per +1 SD of speaker vowel duration or f0.")}
forest_anim <- bind_rows(
  extract_term(m_anim_dur, "vowel_dur_ms_mean_z", "vowel duration"),
  extract_term(m_anim_f0,  "f0_mean_z",           "f0")
) |>
  mutate(model = factor(model, levels = unique(model)))

ggplot(forest_anim, aes(x = value, y = fct_rev(model))) +
  stat_halfeye(.width = c(0.66, 0.95)) +
  geom_vline(xintercept = 0, linetype = 2, colour = "grey40") +
  labs(x = "log-odds shift of \"animate\" judgment per +1 SD acoustic feature",
       y = NULL,
       title = "Perception: which cues drive \"animate\" judgments?")
```

# 9. Discussion and limitations

The data support a partly asymmetric picture of size iconicity. On the production side, both
predicted channels work – speakers reliably lengthen vowels and lower f0 for big referents.
On the perception side only f0 closes the loop: listeners rely on f0 to call "big", but they
do not use vowel duration even though speakers reliably produce it. The duration channel is
production-only in this dataset. This asymmetry is the kind of pattern a pluripotential view of
iconic features would predict (different speakers and different listeners can recruit different
acoustic channels to convey or recover the same semantic dimension), but a directed replication
would be needed to take it as more than suggestive.

Animacy entered the production models as a main effect rather than as a modulator
of the size effect: animate referents in this stimulus set were produced with
shorter vowels and higher f0 regardless of their size, with the size effect itself
essentially unchanged across the animate/inanimate split. We treat this as a
descriptive observation rather than as evidence for animacy qua feature.
A separately-collected per-item cuteness rating (§7.8) provides the cleanest
available diagnostic: when cuteness is added as a covariate to the production
models, the animacy main effect shrinks toward zero, but cuteness itself does
not take its place – both predictors become non-credible. The honest read is
that animals and objects in this stimulus set differ acoustically in ways that
go beyond their size, but that the 36-item design lacks the resolution to
identify which item-level dimension is responsible. Listeners' "animate"
judgments, separately, are modulated by speaker f0 (§7.5), and this perception
pattern survives the cuteness control (§7.8) unchanged.

The exploratory channels are descriptive: CPP and F1 in production, and fricative duration and
fricative SPL in perception, all show credible effects that are worth a closer look in a
study designed around them. We flag them rather than claim them.

Three limitations are worth naming. (i) The stimulus set is small (36 items) and confounds
size with referent identity – speaker-level random effects help but cannot fully partial this
out. (ii) Listeners had a picture in front of them, so item identity is the dominant driver
of any judgment; the acoustic cues modulate on top of an already strong prior (see the
item-level baselines in §6.7). (iii) Animacy is confounded with item-level dimensions of the stimulus set – a
separately-collected cuteness rating (§7.8) is one such dimension, and likely
not the only one – to a degree that the present 36-item design cannot resolve.
A design that orthogonalises animacy and cuteness (and ideally other item-level
dimensions that travel with the animal/object split, such as baseline
expressiveness or articulatory complexity), with more items per cell, would be
needed to identify animacy as a semantic feature in production.

# 10. References {-}

<div id="refs"></div>

# 11. Session info

```{r session}
sessionInfo()
```
